### 2017-06-12 manuscript/manuscript.Rmd
Init empty manuscript.

---
title             : "Curating Research Assets in Behavioral Sciences: A tutorial on managing research materials with R Studio & the Git version control system"
shorttitle        : "Curating Research Assets in Behavioral Sciences"

author: 
  - name: Matti Vuorre
    affiliation: 1
    corresponding: yes
    address: 406 Schermerhorn Hall, 1190 Amsterdam Avenue MC 5501, New York, NY 10027
    email: mv2521@columbia.edu
  - name          : ""
    affiliation   : ""
affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Columbia University"
author_note: >
  Complete departmental affiliations for each author (note the indentation, if you start a new paragraph).

  Enter author note here.

abstract: >
  Enter abstract here (note the indentation, if you start a new paragraph).

keywords          : "keywords"
wordcount         : "X"
bibliography      : ["references.bib"]
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
---

```{r include = FALSE}
library(knitr)  # Make sure that knit directory is project root
library(papaja)
library(tidyverse)
```

\newpage

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

### 2017-06-12 manuscript/manuscript.Rmd
Define reproducibility and distinguish from replicability.

@@
 
+# Introduction
+
+The lack of reproducibility is an increasingly recognized problem across scientific disciplines, and calls for changing the scientific workflow to enhance it have been published in a wide range of research areas, including Biology [@markowetz_five_2015], Ecology [@ihle_striving_2017], Neuroscience [@eglen_toward_2017] and Psychology [@munafo_manifesto_2017].
+However, although exhortations to focus on reproducibility in the scientific practice are now commonplace on the pages of leading scientific journals [@baker_1500_2016], only a small minority of researchers are familiar with the tools and practices that enable implementing reproducibility in the scientific workflow.
+Therefore, although there now is a broad consensus that efforts to improve reproducibility are important, and even on some of the tools that may allow to do so, materials instructing researchers in using them are lacking.
+In this tutorial paper, we present a detailed walk-through of the git version control system for behavioral scientists.
+
+## Reproducibility
+
+Consider the following [@ihle_striving_2017, p.2]: Have you ever found a mistake in your results without knowing what caused it?
+Forgot what analyses you have already done and how (and possibly, why)?
+Lost datasets or information about the cases or variables in a dataset?
+Struggled in redoing analyses when new data became available?
+Had difficulty understanding what data to use or how in a project that you inherited from another researcher?
+Answering any of these questions in the affirmative suggests that your work might benefit from improving reproducibility [@ihle_striving_2017].
+
+But what exactly is reproducibility and why does it matter?
+Reproducibility can be defined as follows: "A research project is computationally reproducible if a second investigator (including you in the future) can recreate the final reported results of the project, including key quantitative findings, tables, and figures, given only a set of files and written instructions." [@kitzes_practice_2017] In the context of experimental Psychology, for example, a project would mean an experiment or a set of experiments investigating a theory or hypothesis, reported results would be a conference presentation or a submitted manuscript.
+In this field, key quantitative findings are usually probability values from statistical models, such as *p*-values, or tables of descriptive statistics such as (differences in) means.
+
+Given this broad definition, here is an example of a non-reproducible project: A published journal article advertises a specific relationship between two variables, to some specified degree of uncertainty, but doesn't provide the raw data or code used to analyse it.
+An example of a reproducible project, on the other hand, provides a well organized package of i) the raw data supporting the claims made in the article, ii) the computer code (or steps of analysis) required to compute the summary and test statistics from the data, and iii) instructions on how to apply ii) to i), if it is not self-evident.
+Clearly, many projects fall in between, and can be partly reproducible--e.g.
+provide raw data but no code.
+
+This definition makes clear an important distinction: reproducibility is not the same as replicability.
+Traditional methods and results sections in journal articles have focused on ensuring replicability (but not reproducibility) by giving detailed instructions on how to repeat the experiment and collect a data set *like* the original one.
+Viewed in this light, replicability is a broad methodological issue concerning the epistemology of the scientific claims; whereas reproducibility--the topic of this tutorial--concerns the minimal steps required to allow checking for the validity of the computations required to assert the scientific claim in the first place.
+
+## Challenges to reproducibility
+
 \newpage

### 2017-06-12 manuscript/manuscript.Rmd
Brief notes on challenges to reproducibility, to be continued.

@@
 
+Although in the abstract reproducibility seems easy, it has in large parts failed in practice.
+One study found that X% of statistical claims in papers could not be reproduced.
+Some reasons for this included the original data not being available; personnel being changed leading to materials not being organized; not clear which data was used for analysis; not clear which analyses were done.
+
+To target these key threats to reproducibility, many authors have offered that research scientists could do better to curate the key assets of their work: the data and code used to analyze the data.
+Recognizing that data in the modern workflow is usually in electronic form, and the procedures to analyze it almost exclusively being run on computers, the empirical sciences have recognized certain aspects of computer science to help with this respect.
+One idea gaining traction in the research community is the idea of version control systems (VCS).
+
 \newpage

### 2017-06-12 manuscript/manuscript.Rmd
Initial thoughts on paper.

@@
 
+## Aims of the current manuscript
+
+We introduce the popular VCS Git.
+Illustrate its use in the scientific workflow with a hypothetical example project.
+Provide commands / instructions for Apple and Windows users.
+Show how to use it with minimal learning curve with the R Studio IDE.
+[TODO]
+
+# The Git version control system
+
+VCSs are computer programs designed for "tracking changes in computer files and coordinating work on those files among multiple people" (Wikipedia).
+In this tutorial, we focus on the most widely used (within science) VCS, called Git.
+It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (??
+Windows), Git is not a standalone program for navigating files and folders on a computer.
+Instead, it adds functionality to the OSs existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI) / integrated development environment (IDE)--that allow taking snapshots of the project and its files, and distributing the work across multiple computers and users.
+To begin using Git, users must first download and install the Git software on their computers.
+Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others), and is used by major software developers such as Microsoft, Google, and Facebook.
+It can be freely downloaded at <https://git-scm.com/>.
+
+## Installing Git
+
+Even if you already have Git installed (some computers do) it is a good idea to install the latest version[^git-version], which can be downloaded as a standalone program from <https://git-scm.com/download>.
+Here, we provide more detailed instructions on installing Git for OS X and Windows, but Linux (and other OS) users can find instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>).
+Along with install instructions, we detail how to configure the Git program so that it is ready to be used.
+
+### OS X
+
+Many OS X computers already have Git installed, especially those operating OS X 10.9 or higher, but it is good practice to install the latest version.
+The easiest way to update Git to the latest version is to download the installer from <http://git-scm.com/download/mac>, and install it like any other program.
+Once Git is installed, it can be used from the command line (the OS X command line is available through the Terminal app, which is included by default with the OS X install), or through various GUIs.
+Although it might be initially intimidating, we encourage using the Terminal app to set the basic configuration of Git.
+
+To operate properly, Git needs to be able to identify its user.
+You can set these by entering a few basic commands in the Terminal.
+To verify the current user information, type `git config --global user.name`, and hit return.
+This should not return anything, unless a previous user of the computer has set it.
+To ensure that Git knows who you are, type `git config --global user.name "User Name"` (where User Name is your name), and hit return.
+If you now re-run the first command, Terminal will return the name you entered as `"User Name"`.
+The second piece of information is your email, which can be entered by `git config --global user.email "email@address.com"` (where email@address.com is your email address).
+You can verify that the correct email address was saved by typing `git config --global user.email` in the Terminal, and hit return.
+Once this information is entered, Git will know who you are, and is thus able to track who is doing what within a project, which is especially helpful when you are collaborating with other people.
+
+### Windows
+
+Windows users can download the Git software installer from <http://git-scm.com/download/win>. [TODO]
+
+# Using Git
+
+The first operating principle of Git is that work is organized into independent projects, which Git calls *repositories*.
+At its core, a repository is a folder on your computer, which is version controlled by Git.
+Everything that happens inside that folder is tracked by Git, but the user has full control of what is tracked (everything, by default) and when.
+To establish the operating principles of Git, we now turn to a practical example using a hypothetical project that, we feel, reflects the components of a typical project in experimental psychology.
+
+## Organizing files and folders
+
+Observers have noted that implementing reproducibility into the scientific workflow is less time-consuming if it is planned from the onset of the project, rather than attempting to add reproducibility to the project after it has been completed [TODO].
+To that end, it is important to organize the project keeping a few key goals in mind (here, we broadly follow established guidelines such as Project TIER recommendations): The files and folders should have easy to understand names (avoid idiosynchratic naming schemes), and the names should indicate the purposes of the files and folders.
+
+The example project has a home directory, which has a concise name indicating its contents.
+Here, we have simply called it `example-project`.
+Everything related to that material is placed in subfolders of the home directory: We have "conducted" one experiment in the project; this experiment had three main components: 1) the computer code that was used to present the stimuli to the participants `experiment-1/experiment/`, 2) the raw data files in `experiment-1/data/`, and 3) the code used to analyse the data in `experiment-1/analysis/`.
+Note that all these folders were sub-sub-directories of the home folder for the project, clearly indicating which experiment (in what may be a multi-experiment project) they refer to.
+
+With this initial set-up in mind, let's begin using Git to enhance the reproducibility of this project.
+
+# Footnotes
+
+[^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
+
 \newpage

### 2017-06-13 manuscript/manuscript.Rmd
Distinguish replicability and reproducibility.

@@
 Viewed in this light, replicability is a broad methodological issue concerning the epistemology of the scientific claims; whereas reproducibility--the topic of this tutorial--concerns the minimal steps required to allow checking for the validity of the computations required to assert the scientific claim in the first place.
+
+Although they are often conlated [TODO] reproducibility and replicability are, in fact, independent of one another.
+One might think that reproducibility [TODO Peng 2011?] is a requirement of replicability, but an experiment can be replicated even if the materials of the original study are not available for replicability.
+Likewise, the materials of a study might allow its results to be reproduced, yet the underlying scientific idea might not be replicable.
+Therefore, it is important to keep these two concepts distinct; in this manuscript we are interested in reproducibility.
 

### 2017-06-13 manuscript/manuscript.Rmd
Work on intro--motivation.

--- 2017-06-12 manuscript/manuscript.Rmd
@@
 
-Although in the abstract reproducibility seems easy, it has in large parts failed in practice.
-One study found that X% of statistical claims in papers could not be reproduced.
-Some reasons for this included the original data not being available; personnel being changed leading to materials not being organized; not clear which data was used for analysis; not clear which analyses were done.
+<!-- TODO work on motivating the use of Git: why is keeping track of materials an issue? Is it an issue? How does VCS help? -->
 
-To target these key threats to reproducibility, many authors have offered that research scientists could do better to curate the key assets of their work: the data and code used to analyze the data.
-Recognizing that data in the modern workflow is usually in electronic form, and the procedures to analyze it almost exclusively being run on computers, the empirical sciences have recognized certain aspects of computer science to help with this respect.
-One idea gaining traction in the research community is the idea of version control systems (VCS).
+Although reproducibility might at first appear easy, it has in large parts failed in practice.
+For example, one survey of leading Psychology journals found that 34-58% of articles published between 1985 and 2013 had at least one inconsistently reported *p*-value [@nuijten_prevalence_2015].
+Upon recalculation, the authors couldn't obtain the same *p*-values from the reported test statistics, and therefore these 34-58% of articles were, to some degree, non-reproducible.
 
-## Aims of the current manuscript
+Why, then, do so many research products fall short of the basic scientific standard of reproducibility?
+We suggest one reason is the poor level of organization and curation of all the research assets (data and code, usually) that play a part in creating the research product.
+Researchers usually have no formal methods or accepted gold standards for curating their research assets, and collaborative workflows, especially, are difficult to manage in a reproducible manner.
+When the complexity and amount of materials related to a project increases, it might be difficult to link analysis files to the correct data files, and track the evolution of both code and data throughout the research cycle.
 
-We introduce the popular VCS Git.
-Illustrate its use in the scientific workflow with a hypothetical example project.
-Provide commands / instructions for Apple and Windows users.
-Show how to use it with minimal learning curve with the R Studio IDE.
-[TODO]
+In fact, the effort to organize and curate materials has been cited as an important reason for why data is so rarely shared in Psychology.
+In one study, researchers asked authors of 141 Psychology articles (with a total of 249 studies) to share their data for reanalysis [@wicherts_poor_2006].
+73% of the authors refused to share their data, and Wicherts et al.
+suggested that they did so because it takes considerable effort to clean, document and organize datasets.
+How could we make organizing and curating materials and data easier, so as to improve the reproducibility of our science?
+
+## Aims of the article
+
+Fortunately for the empirical sciences, challenges related to organizing and curating materials across time, space, and personnel have been solved to a high standard in computer science with Version Control Systems (VCS).
+In the remainder of this article, we introduce a popular VCS called Git, and illustrate its use in the scientific workflow with a hypothetical example project.
+In the tutorial below, we provide show how to use the Git VCS to curate your research materials by using text commands from the computers command line, and with a Graphical User Interface called R Studio.
 
@@
 In this tutorial, we focus on the most widely used (within science) VCS, called Git.
+
+<!-- Insert illustration of VCS about here. -->
+
 It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (??

### 2017-06-13 manuscript/manuscript.Rmd
Changes to 'installing and using Git'

@@
 Here, we provide more detailed instructions on installing Git for OS X and Windows, but Linux (and other OS) users can find instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>).
-Along with install instructions, we detail how to configure the Git program so that it is ready to be used.
+After installation instructions, we detail how to configure the Git program so that it is ready to be used.
 
@@
 Once Git is installed, it can be used from the command line (the OS X command line is available through the Terminal app, which is included by default with the OS X install), or through various GUIs.
-Although it might be initially intimidating, we encourage using the Terminal app to set the basic configuration of Git.
+Although it might at first appear intimidating because of the archaic-looking interface, we encourage using Git from the command line.
 
@@
 You can set these by entering a few basic commands in the Terminal.
+The following commands are intended to be entered to Terminal or a similar command line interface.
 To verify the current user information, type `git config --global user.name`, and hit return.
-This should not return anything, unless a previous user of the computer has set it.
+This should not return anything, unless a previous user of the computer has set the global Git user name.
+
 To ensure that Git knows who you are, type `git config --global user.name "User Name"` (where User Name is your name), and hit return.
@@
 You can verify that the correct email address was saved by typing `git config --global user.email` in the Terminal, and hit return.
-Once this information is entered, Git will know who you are, and is thus able to track who is doing what within a project, which is especially helpful when you are collaborating with other people.
+Once this information is entered, Git will know who you are, and is thus able to track who is doing what within a project, which is especially helpful when you are collaborating with other people, or when you are working on multiple computers.
 
@@
 
-The first operating principle of Git is that work is organized into independent projects, which Git calls *repositories*.
-At its core, a repository is a folder on your computer, which is version controlled by Git.
-Everything that happens inside that folder is tracked by Git, but the user has full control of what is tracked (everything, by default) and when.
-To establish the operating principles of Git, we now turn to a practical example using a hypothetical project that, we feel, reflects the components of a typical project in experimental psychology.
+The first operating principle of Git is that your work is organized into independent projects, which Git calls *repositories*.
+At its core, a repository is a folder on your computer, which is version controlled by Git (you can tell if a folder is "monitored" by git by checking if the folder, or any of its parent folders, contains a hidden `.git` directory).
+Everything that happens inside a repository is tracked by Git, but the user has full control of what is tracked (everything, by default) and when.
+Because the user(s) has full control of what and when is tracked, there is a small set of operations the users need to execute every time they wish to log something in Git.
+
+Briefly, when you work in a Git repository, Git stores the state of each file that it monitors, and when any of these files change, Git indicates that they differ from the previously logged state.
+If the user then is happy with the current changes, she can **add** the changed files to Git's "staging area".
+If the user then is certain that the files added to the staging area are in good shape, they can **commit** the changes.
+These two operations are the backbone of using Git to store the state of the project whenever meaningful changes are made.
+
+Most importantly, each state in Git's log contains the full state of the files at that point in time.
+Users can always go back to an earlier version by "checking out" a previous state of Git's log.
+That's why these programs are called Version Control Systems.
+
+To establish the operating principles of Git in practice, we now turn to a practical example using a hypothetical project that, we feel, reflects the components of a typical project in experimental psychology.
 

### 2017-06-13 manuscript/manuscript.Rmd
How to initialize a git repo.

@@
 To establish the operating principles of Git in practice, we now turn to a practical example using a hypothetical project that, we feel, reflects the components of a typical project in experimental psychology.
+To most clearly show the use of Git, we begin with an empty project with no components.
 
@@
 
-The example project has a home directory, which has a concise name indicating its contents.
-Here, we have simply called it `example-project`.
-Everything related to that material is placed in subfolders of the home directory: We have "conducted" one experiment in the project; this experiment had three main components: 1) the computer code that was used to present the stimuli to the participants `experiment-1/experiment/`, 2) the raw data files in `experiment-1/data/`, and 3) the code used to analyse the data in `experiment-1/analysis/`.
-Note that all these folders were sub-sub-directories of the home folder for the project, clearly indicating which experiment (in what may be a multi-experiment project) they refer to.
+The first step is to create a home folder for the project.
+This folder should have an immediately recognizable name, and should be placed somewhere on your computer where you can find it.
+We call the example project `git-example`.
+All the materials related to this project will be placed in subfolders of the home directory, but for now `git-example` is just an empty folder waiting to be filled with data, code, and documents.
 
-With this initial set-up in mind, let's begin using Git to enhance the reproducibility of this project.
+## Initializing a Git repository
+
+Next, you need to navigate to the folder and initialize it as a Git repository.
+Here, the `git-example` project is in the users `Documents` folder, and we can use `cd ..` in the OS X terminal to navigate up in the folder hierarchy, and `cd <folder>` to navigate into `<folder>`.
+So assuming that the Terminal opens up in `Documents`, we navigate to `git-example` with
+
+```
+cd git-example
+```
+
+Then, to turn this folder into a Git repository, type in
+
+```
+git init
+```
+
+and hit Return.
+This command initializes the folder as a Git repository, and the only change so far has been the addition of a hidden `.git` folder inside `git-example` (and a `.gitattributes` file; but users can ignore these hidden files and folders).
+
+## Adding a file to Git
 

### 2017-06-13 manuscript/manuscript.Rmd
Git basics

@@
 
+Every project, and therefore every repository, should contain a brief not that explains what the project is about and who to contact about it.
+This note is usually called a readme, and therefore our first contribution to this project will be a README file.
+The README file should be a plain text file (i.e.
+not created with Microsoft Word) that can be read with a simple text editor.
+This file can now be seen in the `git-example` folder by using the system file viewer.
+Because we added this file to a Git repository, Git is also aware of it.
+To see what files have changed since the last status change in the repository (there clearly has been only this one), you can ask for Git's **status**:
+
+```bash
+git status
+```
+
+Which in this case would return
+
+```bash
+Matti [~/Documents/git-example]$ git status
+On branch master
+
+Initial commit
+
+Untracked files:
+  (use "git add <file>..." to include in what will be committed)
+
+	README
+
+nothing added to commit but untracked files present (use "git add" to track)
+Matti [~/Documents/git-example]$ 
+```
+
+The relevant output returned from executing this command is the "Untracked files:" part.
+There, Git tells the user that there is an untracked file (README) in the repository.
+To keep track of the status of this file, we **add** it to Git by using the command `git add` followed by either a `.` (for adding all untracked files) or `README` for only tracking the `README` file.
+
+```bash
+git add README
+```
+
+We've now added this file to the staging area, and if we are happy with changes to the file's status (it has been created), we can **commit** the file to Git's versioning system.
+Commits are the most important Git operation: They signify any meaningful change to the repository, and can be later browsed and compared to one another.
+As such, it is helpful to attach a small message to each commit, describing why that commit was made.
+Here, we've created a README file, and our commit command would look as follows:
+
+```bash
+git commit -m "Add README file."
+```
+
+The quoted text after the `-m` flag is the "commit message".
+Entering this command to the command line returns a brief description of the commit, such as how many files changed, and how many characters inside those files were inserted and deleted.
+
+## Keeping track of changes to a file with Git
+
+Most importantly, the `git-example` project now keeps track of all and any changes to README.
+To illustrate, we now add some text to README to describe the project (title, what the project is about, who are involved, who to contact), save the file, and then ask for Git's status with `git status` on the command line:
+
+```bash
+Matti [~/Documents/git-example]$ git status
+On branch master
+Changes not staged for commit:
+  (use "git add <file>..." to update what will be committed)
+  (use "git checkout -- <file>..." to discard changes in working directory)
+
+	modified:   README
+
+no changes added to commit (use "git add" and/or "git commit -a")
+```
+
+Git can now tell that the README file has changed, and we can repeat the add and commit steps to permanently record the current state of the project to Git's log:
+
+```bash
+git add .  # We use the '.' shortcut 
+git commit -m "Populate README with project description."
+```
+
+## What does Git know?
+
+The real importance of these somewhat abstract steps becomes apparent when we consider the Git **log**. We can call
+
+```bash
+git log
+```
+
+To reveal the commit log of this repository.
+The main things to notice in the output of this command are that each commit is associated with a unique hash code (long alphanumeric string), which we can use to call for further information (see below); an author (this is where the earlier setup is apparent); a date and time; and the short commit message.
+Until now, the `git-example` log looks like this:
+
+```bash
+Matti [~/Documents/git-example]$ git log
+commit 60cbe5c9b4a78e500314f791080381030577a035
+Author: Matti Vuorre <mv2521@columbia.edu>
+Date:   Tue Jun 13 17:20:27 2017 -0400
+
+    Populate README with project description.
+
+commit 16c475023ecbc99446164187eeaaab10647ac550
+Author: Matti Vuorre <mv2521@columbia.edu>
+Date:   Tue Jun 13 17:14:14 2017 -0400
+
+    Add README file.
+```
+
+To see what exactly changed in the last commit (the log has latest commits at the top), we can call `git show` with the commit's hash (only relevant parts of output shown below):
+
+```bash
+Matti [~/Documents/git-example]$ git show 60cbe5c9b4a78e500314f791080381030577a035
+commit 60cbe5c9b4a78e500314f791080381030577a035
+Author: Matti Vuorre <mv2521@columbia.edu>
+Date:   Tue Jun 13 17:20:27 2017 -0400
+
+    Populate README with project description.
+
+diff --git a/README b/README
+--- a/README
++++ b/README
+@@ -0,0 +1,8 @@
++# Example Git Project
++This example project illustrates the use of Git.
++authors:
++Matti Vuorre <mv2521@columbia.edu>
++James Curley
++2017
+```
+
+This output can be investigated for a detailed log of all changes created by that commit.
+From top, it lists the author of the commit, the commit message, and then the commit's "**diff**" (i.e.
+what differs in the new version vs the old version of the file.) The current diff shows that 1 file received eight additional lines of text (the part wrapped in @ symbols), and then the additions themselves (the lines prepended with +s).
+
+Although we now understand the fundamentals of using Git to track the states of (and therefore changes to) a repository, this contrived and overly simplistic example doesn't allow full appreciation of the benefits of using Git for version control.
+
 # Footnotes

### 2017-06-13 manuscript/manuscript.Rmd
coauthors!

@@
     email: mv2521@columbia.edu
-  - name          : ""
-    affiliation   : ""
+  - name          : "James P. Curley"
+    affiliation   : "1"
 affiliation:

### 2017-06-14 manuscript/manuscript.Rmd
Improve explanation of what VCS is

@@
 
-# The Git version control system
+# Version Control Systems
 
-VCSs are computer programs designed for "tracking changes in computer files and coordinating work on those files among multiple people" (Wikipedia).
-In this tutorial, we focus on the most widely used (within science) VCS, called Git.
+Version Control Systems (VCS; also known as Source Control systems) are computer programs designed for tracking changes to computer code, and sharing the code---and its history---with others.
+Although initially developed for writing computer code collaboratively, it is easy to recognize the benefits of VCS for scientific production and collaboration as well.
+For example, behavioral experiments are often written in a programming language, and can include multiple authors and versions.
+Keeping track of the versions of the program, and allowing many authors to contribute to it (without breaking the code) are problems that VCSs were specifically designed for.
+By extension, VCS easily adopts to other aspects of the research cycle, such as curating data across computers, laboratories, and experiments; even writing manuscripts is easier when it is recognized that manuscripts (usually) undergo multiple rounds of revision, and that multiple authors might want to keep track of all these different versions.
+
+The core concepts of version control are that contributors to a project create small checkpoints of the changes they make to the source code (or manuscript text, data, etc.), and then submit those changes to the VCS.
+The VCS maintains a history of all these little checkpoints, and the exact state of the project at each of these checkpoints.
 
 <!-- Insert illustration of VCS about here. -->
+
+Some popular Version Control Systems are SVN, Mercurial, and Git.
+In this tutorial, we focus on the most widely used (in science) of these, called Git.
+Git is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.
+
+## The Git Version Control System
 
@@
 It can be freely downloaded at <https://git-scm.com/>.
+Before we explain the basics of using Git, we present brief installation instructions for Windows and Apple users.
 

### 2017-06-14 manuscript/manuscript.Rmd
Add box for Git resources.

@@
 
-Observers have noted that implementing reproducibility into the scientific workflow is less time-consuming if it is planned from the onset of the project, rather than attempting to add reproducibility to the project after it has been completed [TODO].
-To that end, it is important to organize the project keeping a few key goals in mind (here, we broadly follow established guidelines such as Project TIER recommendations): The files and folders should have easy to understand names (avoid idiosynchratic naming schemes), and the names should indicate the purposes of the files and folders.
+Implementing reproducibility into the scientific workflow is less time-consuming if it is planned from the onset of the project, rather than attempting to add reproducibility to the project after it has been completed.
+It is therefore important to organize the project keeping a few key goals in mind (here, we broadly follow established guidelines such as Project TIER recommendations): The files and folders should have easy to understand names (avoid idiosynchratic naming schemes), and the names should indicate the purposes of the files and folders.
 
@@
 Here, the `git-example` project is in the users `Documents` folder, and we can use `cd ..` in the OS X terminal to navigate up in the folder hierarchy, and `cd <folder>` to navigate into `<folder>`.
-So assuming that the Terminal opens up in `Documents`, we navigate to `git-example` with
+So assuming that the Terminal opens up in `Documents`, we navigate to `git-example` with (to execute code in Terminal, enter the text into the command line and hit Return):
 
@@
 
-Then, to turn this folder into a Git repository, type in
+Then, to turn this folder into a Git repository, execute
 
@@
 
-and hit Return.
 This command initializes the folder as a Git repository, and the only change so far has been the addition of a hidden `.git` folder inside `git-example` (and a `.gitattributes` file; but users can ignore these hidden files and folders).
@@
 
+# Further resources
+
+<!-- Text box: Resources for learning Git -->
+
+\noindent
+\fbox{\begin{minipage}{\textwidth}
+  \begin{itemize}
+    \item \href{https://www.git-tower.com/blog/workflow-of-version-control}{Basic VCS workflow}, an infographic explaining how VCS works.
+    \item \href{https://services.github.com/resources/cheatsheets/}{GitHub's Git cheat sheet} is available in multiple languages and contains the most used Git commands.
+    \item \href{https://try.github.io/levels/1/challenges/1}{TryGit}, an interactive website for learning the basics of Git.
+    \item \href{http://r-pkgs.had.co.nz/git.html}{Git + GitHub} in an R programming context.
+    \item \href{https://git-scm.com/book/en/v2}{Pro Git book}, a complete manual of Git.
+  \end{itemize}
+\end{minipage}}
+
 # Footnotes

### 2017-06-14 manuscript/manuscript.Rmd
Add table of main Git commands

@@
 Although we now understand the fundamentals of using Git to track the states of (and therefore changes to) a repository, this contrived and overly simplistic example doesn't allow full appreciation of the benefits of using Git for version control.
+To better illustrate Git's functioning, we now fast-forward in the hypothetical example project to a stage in the project where more files and materials have been created.
+
+## (Slightly more) advanced Git
+
+## Overview of the Git Workflow
+
+```{r, results = 'asis', warning = F}
+commands <- read_csv("data/git-commands.csv")
+commands %>% 
+    apa_table(
+        caption = "Main Git commands.",
+        note = "Source: https://help.github.com/articles/github-glossary/ and https://services.github.com/on-demand/downloads/github-git-cheat-sheet.pdf",
+        landscape = FALSE)
+```
 

### 2017-06-14 manuscript/manuscript.Rmd
Explain .gitignore

@@
 
-```
+```bash
 cd git-example
@@
 
-```
+```bash
 git init
@@
 On branch master
-
 Initial commit
-
 Untracked files:
   (use "git add <file>..." to include in what will be committed)
-
 	README
-
 nothing added to commit but untracked files present (use "git add" to track)
@@
 
+The state of the project now looks like this:
+
+```bash
+git-example/
+    admin/
+        ethics-info.pdf
+    experiment-1/
+        analysis/
+            plan_n.R
+    manuscript/
+    README
+```
+
+Running `git status` now tells that there are two new files (empty folders are ignored).
+That is, since the previous commit, two files have been added to the project; one a pdf with some administrative information (`ethics-info.pdf`), the other one is an R script for a pre-planned power analysis (`plan_n.R`).
+We would certainly like to track any changes to the power analysis script, but the ethics information file isn't really something that we need to keep track of---we can't make changes to it anyway.
+It would be laborious to constantly keep ignoring it when committing changes to Git (especially if we'd like to commit multiple files simultaneously with `git add .`) Git has an elegant solution to specifying which files to keep track of.
+Because by default all files are tracked, Git uses a special file for instructing which files are to be *ignored* from tracking.
+
+### Make Git Ignore Files
+
+To make Git ignore files, we simply add a plain text file called `.gitignore` to the home folder of the repository.
+Each row of this file should specify a file or a folder (or a regular expression) that Git should ignore.
+In our example we want to make Git ignore the `admin/` folder entirely, and any file with the .pdf extension inside `manuscript/.
+The example .gitignore file looks like this (lines beginning with `#` are comments):
+
+```bash
+# Ignore everything inside the 'admin' folder
+admin/
+
+# Ignore all .pdf files in the 'manuscript' folder
+manuscript/*.pdf
+```
+
+Re-running `git status` now only shows the `plan_n.R` file (and the newly created .gitignore file, which is also under version control, naturally.)
+
+Because there are now two untracked files, and we would like to keep a clean history of what has happened in the project, we create two separate commits: One for the .gitignore file, and one for the power analysis file.
+
+```bash
+git add .gitignore
+git commit -m "Added .gitignore file"
+
+git add .
+git commit -m "Completed power analysis"
+```
+
+After this last commit, if we ever want to know in the future what our planned sample size was, we can come back to this commit with `git log` or `git show` and see what was inside the newly created power analysis file.
+This possibility of "rewinding history" is especially useful for files that might go under multiple revisions, or if we might be interested when and in what order the files were created.
+One might consider committing a power analysis as a small personal pre-registration of a part of the research plan.
+
+# Git Summary
+
 ## Overview of the Git Workflow
@@
 
-# Further resources
-
-<!-- Text box: Resources for learning Git -->
+## Further resources for learning Git
 

### 2017-06-14 manuscript/manuscript.Rmd
Explain .gitignore
Introduce collaborating w/ centralized workflow

@@
 After this last commit, if we ever want to know in the future what our planned sample size was, we can come back to this commit with `git log` or `git show` and see what was inside the newly created power analysis file.
+For instance, if new information suggests that we should change assumed effect size in the power analysis, we can simply edit and save the file, then add and commit the changes to Git with a helpful message that logs this important event in the Git memory of the repository.
+
 This possibility of "rewinding history" is especially useful for files that might go under multiple revisions, or if we might be interested when and in what order the files were created.
 One might consider committing a power analysis as a small personal pre-registration of a part of the research plan.
+
+# Collaborating
+
+The true advantages of using a VCS such as Git become apparent when we consider projects with more than one contributor.
+For example, consider a project where data is collected at multiple sites, and these files are then saved onto a centralized server, or shared through a service that automatically merges files from multiple sources (such as the popular Dropbox service).
+If two or more sites accidentally save a data file with the same name (e.g.
+`data-001.csv`), and these changes are automatically merged, the later file will simply overwrite the earlier file.
+Disaster!
+
+Or consider a data analysis where two or more people work simultaneously on some complicated analysis.
+If user A and B have the same file open on multiple computers, after a while when user B saves the file, user A's version of the file will be overwritten.
+Disaster!
+
+Git and other VCSs were specifically designed to allow (and facilitate) multi-site collaboration on extremely complex projects, as for example required by leading software companies such as Apple and Microsoft.
+Next, we'll illustrate how collaborating on a Git project works in practice.
+
+There are many ways in which a team could collaborate on a Git project (e.g.
+<https://www.atlassian.com/git/tutorials/comparing-workflows>), and here we focus on a common one, called the "Centralized Workflow".
+
+## Overview of the Centralized Git Workflow
+
+In this workflow, a central repository hosted on a research team's server, or an online service like GitHub (<https://github.com/>), is used to pull and push (see Table \@ref(tab:git-commands)) information to and from local repositories.
+First, one user creates this central repository, then other users clone a local copy (see Table \@ref(tab:git-commands)).
+All changes are first created, added, and committed locally as detailed above, and only then **pushed** to the central repository.
+To get changes that other users have pushed to the central repository, a user **pulls** changes from it.
+
+Setting up a centralized Git repository on a research team's private server is relatively straightforward, but because the details vary from team to team, here we illustrate the Centralized Workflow using GitHub.
+
+## GitHub
+
+GitHub is one of the 100 most popular website worldwide, and hosts over 60 million software projects with a total of over 20 million users (<https://github.com/about>).
+We chose GitHub for collaboration because it is already the de-facto standard in VCS collaboration among scientists, it offers free private (see below) repositories for students, and has many attractive features, some of which we will detail below.
 
@@
 
-```{r, results = 'asis', warning = F}
+```{r git-commands, results = 'asis', warning = F}
 commands <- read_csv("data/git-commands.csv")

### 2017-06-14 manuscript/manuscript.Rmd
Create a github repo

@@
 
+Anyone wanting to use GitHub must first create a free user account at <https://github.com>.
+After the user account has been set up, and the user has logged in, creating a new repository at GitHub is self evident (there is a big green button labeled "New repository").
+
+### Create a new GitHub repository
+
+Although we are creating a new repository here, the goal is to create an online "remote" for the running example within this tutorial, and therefore the specific steps are slightly different than what they would be if a brand new repository was needed.
+
+For the current example, after clicking "New repository", we simply enter a name, which can be anything, but for consistency we call the GitHub repository `git-example`.
+After entering the name, click "Create repository", and the next step is to link the new remote to the local repository.
+The required steps are detailed on screen when you do this at GitHub, but here we would execute the following command in the command line:
+
+```bash
+git remote add origin https://github.com/<username>/<reponame>.git
+```
+
+Where `<username>` and `<reponame>` should be replaced with the user's GitHub username, and the GitHub repository name.
+Once the connection is set up, we can **push** local changes to the GitHub remote:
+
+```bash
+git push -u origin master
+```
+
+The `-u origin master` are only required for the first push, as they set up the connection. For pushing any changes collowing this, simply type `git push` after adding and committing locally.
+
+The team has now created the remote Central repository, and other users can start contributing to it.
+
 # Git Summary

### 2017-06-14 manuscript/manuscript.Rmd
Contributing to central repo

@@
 
+### Contributing to a Central (GitHub) repository
+
+Once another team member has set up Git on their own computer, and signed up for GitHub, they need to **clone** the remote GitHub repository onto their local computer.
+[TODO actually check that these steps are correct with a 'fresh' computer.]
+
+To clone a repository, the new user must first navigate to an appropriate location on the computer where they would like to create the repository on their computer, for example their `username/Documents` folder.
+Once the appropriate location is found, cloning will create a new folder for the repo inside this folder.
+
+```bash
+git clone https://github.com/<username>/<reponame>.git
+```
+
+These new contributors can then work on their local copies as detailed in earlier parts of this tutorial; making changes, adding, and committing.
+After committing their changes, they can update the status of the Central Git repo by pushing their changes to it:
+
+```bash
+git push
+```
+
 # Git Summary

### 2017-06-14 manuscript/manuscript.Rmd
Pull and conflicts.

@@
 
+### Obtaining other's changes from Central repo
+
+Just as you must manually push your own local changes to the remote repository, you must also obtain others' changes by **pulling** them from the central repo.
+Before starting to work on your proposed change, pull the remote changes with [TODO see if it needs to be rebased]:
+
+```bash
+git pull
+```
+
+The way in which users, and their local repositories, interact with the central repository by pushing and pulling is the cornerstone of collaboration on GitHub, and thoughtful use of these commands allows for complex workflows without any important code (data, ideas in manuscript, analysis code) being ever overwritten.
+However, there is no automatic way for a computer to tell what changes to prioritize: If two users have worked on the same code, and they both attempt to push their changes, the user to do this later will have to resolve the **conflict**.
+
+### Resolving a conflict
+
+If user A has pushed changes to the central repository while user B was working on the same code, user B's push will result in a merge conflict.
+That is just a natural consequence of two individuals working simultaneously on the same idea.
+User B simply needs to first merge the changes in the remote repository to her own code by pulling from the remote.
+Once the conflict has been manually resolved (and user B's code updated with the pull), user B can push changes from her local repository to the central repository.
+How these potential conflicts appear depends on how users collaborate with one another, and a detailed explanation of all potential scenarios is outside the scope of this tutorial.
+For more information, see e.g.
+<https://www.atlassian.com/git/tutorials/comparing-workflows#centralized-workflow>.
+The GitHub customer service is also very responsive to their users help requests.
+
 # Git Summary

### 2017-06-14 manuscript/manuscript.Rmd
GitHub: private versus public repos

@@
 
+## Private or public collaboration?
+
+By default, all GitHub repositories are public, meaning that anyone with an internet connection can use their web browser to inspect the contents of your repository, or even clone it to their computer.
+This may sound intimidating to researchers who are used to working in a more private context, and clearly necessitates planning and thought with respect to issues such as data privacy and sharing sensitive materials.
+However, for many projects---including the writing of this tutorial---we see very few downsides to working "in the open".
+
+There are two alternatives to working in a public GitHub repository: One, which we won't cover here, is to not use GitHub but instead place the central repository on the research team's shared server.
+The second option is to change the repository's settings on GitHub (this can be done when the repository is first created or afterwards) to private.
+Private repositories, and their contents, are only accessible to invited team members, and are therefore ideal for small teams who would like to work without revealing their master plans to the public just yet.
+
+To make a GitHub repository private, navigate to the repository's website with a web browser (e.g.
+<https://github.com/<username>/<reponame>>), and click "Settings", then "Make this repository private".
+Once one user has set the central GitHub repository to as private, anyone wishing to clone, push, pull, or view the repo must provide their GitHub username and password.
+Only if they match an invited team members username and password can the user access the repository.
+
 # Git Summary

### 2017-06-14 manuscript/manuscript.Rmd
Introduce R Studio for Git from R Studio tutorial

@@
 ---
-title             : "Curating Research Assets in Behavioral Sciences: A tutorial on managing research materials with R Studio & the Git version control system"
+title             : "Curating Research Assets in Behavioral Sciences: A tutorial on the Git version control system"
 shorttitle        : "Curating Research Assets in Behavioral Sciences"
@@
 For more information, see e.g.
-<https://www.atlassian.com/git/tutorials/comparing-workflows#centralized-workflow>.
-The GitHub customer service is also very responsive to their users help requests.
+<https://www.atlassian.com/git/tutorials/comparing-workflows>.
+The GitHub customer service is also very responsive to users' help requests.
 
@@
 Only if they match an invited team members username and password can the user access the repository.
+
+# Using Git from IDE
+
+Finally, we have above detailed how to use Git from the computer's command line, but not all users are comfortable---although there is no reason not to be!---with this mode of interacting with their computers.
+We chose to introduce Git from the command line because eventually users will need to use it for more complicated operations (and it is required for setting the user's information); however, there are various graphical user interfaces (GUIs) available for Git, that allow interacting with Git by pointing and clicking buttons with the mouse.
+
+In this section, we introduce how to use the popular R Studio Integrated Development Environment (IDE) for the R programming language for interacting with Git.
+An IDE is an interface that bundles together many necessary features of software development---some users may be familiar with R Studio for conducting statistical analyses.
+Version control is an essential feature of software development, and R Studio provides a good GUI within its IDE for controlling Git.
+
+For this example, we will begin a completely new repository using only R Studio, and assume that readers are familiar with the concepts from the above tutorial.
+
+## R Studio
+
+[TODO download and install R Studio, cite]
+
+Managing projects (and Git repositories) with R Studio is centered on the idea of R Projects.
+To start a new project with R Studio, open the R Studio application, and click File -> New Project.
+This brings up a dialog asking whether to create a project in a new directory, existing directory, or checkout an existing project from a version control repository.
+Here, we create a new project in a new directory, and choose "New Project" in the following screen.
+Then we'll give a name to the project's home folder and choose where to save it on the computer.
+Importantly, we'll also check the "Create a git repository" box, which will automatically set up a new repository for the project (provided you have set Git up as detailed above).
+Clicking "Create project" creates the folder in the specified location, and two files inside the project's main folder.
+
+[TODO put some screenshots in here]
+
+One of these files is `.gitignore` which we covered above.
+The other file is and .Rproj file, which indicates that the folder is the home folder for an R (Studio) project.
+Users don't interact with this file, but when opened, it is simply a plain text file containing the project's settings (these can be modified through Tools -> Project Options in R Studio).
 
@@
 
-# Footnotes
+\newpage
 
 [^git-version]: As of the writing of this article, the current version of Git is 2.13.1. 
-
-\newpage
-
 # References

### 2017-06-14 manuscript/manuscript.Rmd
RStudio: add, commit changes, add remote, images

@@
 
+## Using Git with R Studio
+
+Once the R Project has been created, R Studio has a "Git" tab in the top-right panel of the GUI.
+At first, this tab shows the two new files in the repository, and some buttons with familiar looking names ("Commit", "Push", etc.)
+
+```{r rstudio-git-1, fig.cap = "R Studio's Git tab for a newly created project."}
+knitr::include_graphics("data/rstudio-git-1.png")
+```
+
+### Add and Commit changes
+
+To mark this milestone of creating a project, let's commit these changes to Git.
+The workflow is exactly the same as when using Git from the terminal, but we now have some visually appealing helpers.
+To begin committing these changes, click on "Commit" in the Git tab.
+This brings up another window where we can select files to add into the staging area (the `.gitignore` file in Figure \@ref(fig:rstudio-git-2) is staged by checking the box).
+Once all the desired files (here we chose both) are staged, we write a short commit message as detailed above, and click the "Commit" button.
+
+```{r rstudio-git-2, fig.cap = "Creating a Git commit with R Studio."}
+knitr::include_graphics("data/rstudio-git-2.png")
+```
+
+### Add a remote GitHub repository
+
+When creating the R Project, there was an option for creating the project from an existing Git repository.
+Because we didn't do this, we must now manually instruct Git that this repo should have a remote in GitHub.
+This is done exactly as above, and we must use command line functions.
+After you've created a new repository on GitHub, navigate to the project's folder on your computer with the command line navigator of your choice, and execute the following commands:
+
+```bash
+git remote add origin https://github.com/<username>/<reponame>.git
+git push -u origin master
+```
+
+After executing these commands, you can use all Git and GitHub features from R Studio without having to use the command line.
+
+### Further examples
+
 # Git Summary
 
-## Overview of the Git Workflow
+## Overview of Git Commands
 

### 2017-06-14 manuscript/manuscript.Rmd
OK so it's impossible to knitr::include_images from other folders. Shame.

@@
 ```{r rstudio-git-1, fig.cap = "R Studio's Git tab for a newly created project."}
-knitr::include_graphics("data/rstudio-git-1.png")
+knitr::include_graphics("images/rstudio-git-1.png")
 ```
@@
 ```{r rstudio-git-2, fig.cap = "Creating a Git commit with R Studio."}
-knitr::include_graphics("data/rstudio-git-2.png")
+knitr::include_graphics("images/rstudio-git-2.png")
 ```
@@
 [^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
+
 # References

### 2017-06-14 manuscript/manuscript.Rmd
Create a table of specific reproducibility challenges addressed by Git

@@
 Git is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.
+
+<!-- Examples of problems in psychology solved by VCS somewhere here. -->
+
+\noindent
+\fbox{\begin{minipage}{\textwidth}
+  \begin{itemize}
+    \item Problem: Researchers usually want to try different ways of visualizing and modeling their data, and sometimes it is difficult to keep track of all the different versions.
+      \begin{itemize}
+        \item Solution: VCS allows researchers to save each version of the analysis into a history from where it can be easily recalled.
+This approach also eschews the need for creating multiple copies of the same analysis files with small tweaks, as illustrated \href{http://phdcomics.com/comics/archive.php?comicid=1531}{here}.
+      \end{itemize}
+  \end{itemize}
+  \begin{itemize}
+    \item Problem:
+      \begin{itemize}
+        \item Solution:
+      \end{itemize}
+  \end{itemize}
+\end{minipage}}
 

### 2017-06-15 manuscript/manuscript.Rmd
Fill YAML info

@@
 ---
-title             : "Curating Research Assets in Behavioral Sciences: A tutorial on the Git version control system"
+title             : "Curating Research Assets in Behavioral Sciences: A Tutorial on the Git Version Control System"
 shorttitle        : "Curating Research Assets in Behavioral Sciences"
@@
   - id            : "1"
-    institution   : "Department of Psychology, Columbia University"
+    institution   : "Department of Psychology, Columbia University, New York, USA"
 author_note: >
@@
 abstract: >
-  Enter abstract here (note the indentation, if you start a new paragraph).
+  Recent calls for improving the reproducibility of research products in behavioral sciences have called for increased attention to the way in which researchers curate, collaborate, and share their research assets, but have failed to provide practical instructions on how to do so.
+In this tutorial paper, we explain how version control systems, such as the popular Git program, are especially suitable for the challenges to reproducibility faced by behavioral science regarding research asset curating, sharing, and collaborating.
+We then present a complete tutorial on how to use Git from the computer's command line, and with the popular graphical interface contained in the R Studio development environment.
+Git is easy to learn, presents an elegant solution to problems related to curating research assets, facilitates multi-site collaboration and productivity (by allowing multiple collaborators to work on the same source simultaneously) and can be implemented on common workflows with minimal extra effort.
+Git may also offer a suitable solution to transparent data (and material) sharing through popular online services, such as GitHub and Open Science Framework.
 
-keywords          : "keywords"
+keywords          : "reproducibility; open science; version control; git; research methods"
 wordcount         : "X"
@@
 lang              : "english"
-class             : "man"
+class             : "doc, 12pt, draftfirst"  # man / jou / doc
 output            : papaja::apa6_pdf
@@
 library(knitr)  # Make sure that knit directory is project root
+opts_chunk$set(
+    fig.align = "center"
+)
 library(papaja)

### 2017-06-15 manuscript/manuscript.Rmd
Remove author note.

@@
     institution   : "Department of Psychology, Columbia University, New York, USA"
-author_note: >
-  Complete departmental affiliations for each author (note the indentation, if you start a new paragraph).
-
-  Enter author note here.
 

### 2017-06-15 manuscript/manuscript.Rmd
Clarify Git & GitHub configuration

@@
 
+### Windows
+
+Windows users can download the Git software installer from <http://git-scm.com/download/win>. [TODO]
+
+## Git setup
+
 To operate properly, Git needs to be able to identify its user.
-You can set these by entering a few basic commands in the Terminal.
-The following commands are intended to be entered to Terminal or a similar command line interface.
+You can set these by entering a few basic commands in the command line.
 To verify the current user information, type `git config --global user.name`, and hit return.
@@
 
-To ensure that Git knows who you are, type `git config --global user.name "User Name"` (where User Name is your name), and hit return.
-If you now re-run the first command, Terminal will return the name you entered as `"User Name"`.
+To ensure that Git knows who you are, type `git config --global user.name "User Name"` (where User Name is your name) in the command line, and hit return.
+If you now re-run the first command, the command line will return the name you entered as `"User Name"`.
+The username can be anything you'd like, but it is probably a good idea to use your real name so that potential collaborators know who you are.
 The second piece of information is your email, which can be entered by `git config --global user.email "email@address.com"` (where email@address.com is your email address).
@@
 Once this information is entered, Git will know who you are, and is thus able to track who is doing what within a project, which is especially helpful when you are collaborating with other people, or when you are working on multiple computers.
-
-### Windows
-
-Windows users can download the Git software installer from <http://git-scm.com/download/win>. [TODO]
 
@@
 After the user account has been set up, and the user has logged in, creating a new repository at GitHub is self evident (there is a big green button labeled "New repository").
+You must create the GitHub account with the same email address as you used above when configuring your local Git (or re-configure your local Git to use your the email address that you used to register for GitHub).
 

### 2017-06-15 manuscript/manuscript.Rmd
Draft 1 of # Introduction

@@
 abstract: >
-  Recent calls for improving the reproducibility of research products in behavioral sciences have called for increased attention to the way in which researchers curate, collaborate, and share their research assets, but have failed to provide practical instructions on how to do so.
-In this tutorial paper, we explain how version control systems, such as the popular Git program, are especially suitable for the challenges to reproducibility faced by behavioral science regarding research asset curating, sharing, and collaborating.
-We then present a complete tutorial on how to use Git from the computer's command line, and with the popular graphical interface contained in the R Studio development environment.
-Git is easy to learn, presents an elegant solution to problems related to curating research assets, facilitates multi-site collaboration and productivity (by allowing multiple collaborators to work on the same source simultaneously) and can be implemented on common workflows with minimal extra effort.
+  Recent calls for improving the reproducibility of behavioral sciences have increased attention to the ways in which researchers curate, share and collaborate on their research assets.
+However, these discussions have largely failed to provide practical instructions on how to do so.
+In this tutorial paper, we explain why version control systems, such as the popular Git program, are especially suitable for these challenges to reproducibility.
+We then present a tutorial on how to use Git from the computer's command line, and with the popular graphical interface contained in the R Studio development environment.
+This tutorial is especially written for behavioral scientists with no previous experience with version control systems.
+Git is easy to learn, presents an elegant solution to challenges to reproducibility related to curating research assets, facilitates multi-site collaboration and productivity (by allowing multiple collaborators to work on the same source simultaneously) and can be implemented on common behavioral science workflows with little extra effort.
 Git may also offer a suitable solution to transparent data (and material) sharing through popular online services, such as GitHub and Open Science Framework.
 
-keywords          : "reproducibility; open science; version control; git; research methods"
+keywords          : "reproducibility; version control; git; research methods; open science"
 wordcount         : "X"
@@
 
-The lack of reproducibility is an increasingly recognized problem across scientific disciplines, and calls for changing the scientific workflow to enhance it have been published in a wide range of research areas, including Biology [@markowetz_five_2015], Ecology [@ihle_striving_2017], Neuroscience [@eglen_toward_2017] and Psychology [@munafo_manifesto_2017].
-However, although exhortations to focus on reproducibility in the scientific practice are now commonplace on the pages of leading scientific journals [@baker_1500_2016], only a small minority of researchers are familiar with the tools and practices that enable implementing reproducibility in the scientific workflow.
-Therefore, although there now is a broad consensus that efforts to improve reproducibility are important, and even on some of the tools that may allow to do so, materials instructing researchers in using them are lacking.
-In this tutorial paper, we present a detailed walk-through of the git version control system for behavioral scientists.
+The lack of reproducibility is increasingly recognized as a problem across scientific disciplines, and calls for changing the scientific workflow to enhance reproducibility have been published in a wide range of research areas, including Biology [@markowetz_five_2015], Ecology [@ihle_striving_2017], Neuroscience [@eglen_toward_2017] and Psychology [@munafo_manifesto_2017].
+However, although calls to focus on reproducibility are now commonplace on the pages of leading scientific journals [e.g.
+@baker_1500_2016], only a small minority of researchers are familiar with the tools and practices that enable implementing reproducibility in their workflows.
+Therefore, although there now is a broad consensus that efforts to improve reproducibility are important, materials instructing researchers in using them are lacking.
 
@@
 
-Consider the following [@ihle_striving_2017, p.2]: Have you ever found a mistake in your results without knowing what caused it?
+Consider for a while if you have ever struggled with the following questions [@ihle_striving_2017, p.2]: Have you ever found a mistake in your results without knowing what caused it?
 Forgot what analyses you have already done and how (and possibly, why)?
 Lost datasets or information about the cases or variables in a dataset?
-Struggled in redoing analyses when new data became available?
+Struggled in redoing statistical or computational analyses when new data became available?
 Had difficulty understanding what data to use or how in a project that you inherited from another researcher?
@@
 
-But what exactly is reproducibility and why does it matter?
-Reproducibility can be defined as follows: "A research project is computationally reproducible if a second investigator (including you in the future) can recreate the final reported results of the project, including key quantitative findings, tables, and figures, given only a set of files and written instructions." [@kitzes_practice_2017] In the context of experimental Psychology, for example, a project would mean an experiment or a set of experiments investigating a theory or hypothesis, reported results would be a conference presentation or a submitted manuscript.
-In this field, key quantitative findings are usually probability values from statistical models, such as *p*-values, or tables of descriptive statistics such as (differences in) means.
+But what exactly is reproducibility?
+Reproducibility can be defined as follows: "A research project is computationally reproducible if a second investigator (including you in the future) can recreate the final reported results of the project, including key quantitative findings, tables, and figures, given only a set of files and written instructions." [@kitzes_practice_2017].
+In the context of experimental Psychology, for example, a "project" could be an experiment or a set of experiments investigating a theory or hypothesis, "reported results" could be figures or statistical analyses reported in a conference presentation or a manuscript.
 
-Given this broad definition, here is an example of a non-reproducible project: A published journal article advertises a specific relationship between two variables, to some specified degree of uncertainty, but doesn't provide the raw data or code used to analyse it.
-An example of a reproducible project, on the other hand, provides a well organized package of i) the raw data supporting the claims made in the article, ii) the computer code (or steps of analysis) required to compute the summary and test statistics from the data, and iii) instructions on how to apply ii) to i), if it is not self-evident.
-Clearly, many projects fall in between, and can be partly reproducible--e.g.
-provide raw data but no code.
+This definition makes an important distinction between *reproducibility* and *replicability*.
+Methods and results sections in traditional journal articles have focused on ensuring replicability by giving detailed instructions on how to repeat the experiment and collect a data set similar to the original one.
+However, little or no emphasis is placed on issues related to reproducibility, such as details of the data processing pipeline and statistical model exploration, although these issues are often more important to the reliability of the scientific results than are their end results (e.g.
+reported *p*-values; @leek_statistics:_2015).
+To put it more generally, replicability is a broad conceptual issue concerning the epistemology of scientific claims.
+Reproducibility, on the other hand, is a technical requirement for allowing others (including researchers themselves) to assess the reliability of a specific set of empirical or computational results.
 
-This definition makes clear an important distinction: reproducibility is not the same as replicability.
-Traditional methods and results sections in journal articles have focused on ensuring replicability (but not reproducibility) by giving detailed instructions on how to repeat the experiment and collect a data set *like* the original one.
-Viewed in this light, replicability is a broad methodological issue concerning the epistemology of the scientific claims; whereas reproducibility--the topic of this tutorial--concerns the minimal steps required to allow checking for the validity of the computations required to assert the scientific claim in the first place.
+Further, although they are often erroneously conlated [@open_science_collaboration_estimating_2015; @peng_reproducible_2011] reproducibility and replicability are independent of one another.
+Although the distinction between the two may be somewhat blurred in purely computational areas (e.g.
+simulation studies; @peng_reproducible_2011), it is clear that an experiment can be replicated even if the materials of the original study are not available for replicability.
+Likewise, the materials of an experiment might allow its results to be reproduced, yet the underlying scientific hypothesis might not be replicable.
+Therefore, it is important to keep these two concepts distinct; in the remainder of the manuscript we explicitly only discuss reproducibility.
 
-Although they are often conlated [TODO] reproducibility and replicability are, in fact, independent of one another.
-One might think that reproducibility [TODO Peng 2011?] is a requirement of replicability, but an experiment can be replicated even if the materials of the original study are not available for replicability.
-Likewise, the materials of a study might allow its results to be reproduced, yet the underlying scientific idea might not be replicable.
-Therefore, it is important to keep these two concepts distinct; in this manuscript we are interested in reproducibility.
+Given this broad definition of reproducibility, it is easy to recognize that many projects fall between the extremes of completely replicable and not replicable at all.
+A completely reproducible project would provide consumers the entire raw data set, and detailed instructions to analyze it (usually as unambigous computer code).
+This degree of reproducibility is very rare in Psychology [@wicherts_poor_2006; @vanpaemel_are_2015], although data sharing has become more common in recent years in journals that award authors who do so with special badges [@kidwell_badges_2016].
+However, even when authors (report to) share data, the shared data sets sometimes turn out to be incorrect, not usable, incomplete, or not available at all [@kidwell_badges_2016].
+This suggests that reproducibility requires more than good intentions.
+What, then, are some specific challenges to reproducibility?
 
@@
 
-<!-- TODO work on motivating the use of Git: why is keeping track of materials an issue? Is it an issue? How does VCS help? -->
+Evidence suggests that the level of reproducibility within Psychology is not very high.
+One survey of leading Psychology journals found that 34-58% of articles published between 1985 and 2013 had at least one inconsistently reported *p*-value [@nuijten_prevalence_2015]: Upon recalculation, the authors couldn't obtain the same *p*-values from the reported test statistics, and therefore these 34-58% of articles were, to some degree, non-reproducible (the inferential statistics could not be reproduced).
 
-Although reproducibility might at first appear easy, it has in large parts failed in practice.
-For example, one survey of leading Psychology journals found that 34-58% of articles published between 1985 and 2013 had at least one inconsistently reported *p*-value [@nuijten_prevalence_2015].
-Upon recalculation, the authors couldn't obtain the same *p*-values from the reported test statistics, and therefore these 34-58% of articles were, to some degree, non-reproducible.
+Why, do so many research products fall short of the basic scientific standard of reproducibility?
+We suggest one reason is the poor level at which researchers organize, curate, and collaborate on the research assets---e.g.
+data and code used to analyze it---which support the product's conclusions.
+In one study, researchers asked authors of 141 Psychology articles (with a total of 249 studies) to share their data for reanalysis.
+73% of the authors refused to share their data, leading the study's authors to suggest that the considerable efforts involved in cleaning datasets at the final stages of publication may make data sharing unattractive to many [@wicherts_poor_2006; see also @vanpaemel_are_2015].
+Vanpaemel et al.
+highlight some reasons provided by authors for not sharing their data:
 
-Why, then, do so many research products fall short of the basic scientific standard of reproducibility?
-We suggest one reason is the poor level of organization and curation of all the research assets (data and code, usually) that play a part in creating the research product.
-Researchers usually have no formal methods or accepted gold standards for curating their research assets, and collaborative workflows, especially, are difficult to manage in a reproducible manner.
-When the complexity and amount of materials related to a project increases, it might be difficult to link analysis files to the correct data files, and track the evolution of both code and data throughout the research cycle.
+>"To our surprise, some authors are apparently willing to share, but have no easy access to their own data or have lost their data altogether, due to computer crashes or collaborators having left the university.
+Many authors cite a lack of time as a reason not to share, and note that sharing their data would take too much effort, which is probably due to poor documenta- tion and storage practices." (p.
+2-3)
 
-In fact, the effort to organize and curate materials has been cited as an important reason for why data is so rarely shared in Psychology.
-In one study, researchers asked authors of 141 Psychology articles (with a total of 249 studies) to share their data for reanalysis [@wicherts_poor_2006].
-73% of the authors refused to share their data, and Wicherts et al.
-suggested that they did so because it takes considerable effort to clean, document and organize datasets.
-How could we make organizing and curating materials and data easier, so as to improve the reproducibility of our science?
+In short, scientists usually excel at *producing* research assets, but commonly fail at *curating* them (a topic commonly known as Digital Asset Management; @jacobsen_implementing_2012).
+This state of affairs is not surprising because researchers are trained to do research, but receive little or no training in formal methods or accepted gold standards for curating their research assets.
+For simple[^simple] projects, such as writing a solo-author manuscript, this has not been a problem.
+However, with increasing complexity of research designs, number of collaborators, research sites, and technical skills required in modern scientific workflows, not knowing how to keep track of materials can lead to serious problems, as detailed above.
 
-## Aims of the article
-
-Fortunately for the empirical sciences, challenges related to organizing and curating materials across time, space, and personnel have been solved to a high standard in computer science with Version Control Systems (VCS).
+Fortunately for the empirical sciences, challenges related to organizing and curating materials ("digital assets") across time, space, and personnel have been solved to a high standard in computer science with Version Control Systems (VCS).
 In the remainder of this article, we introduce a popular VCS called Git, and illustrate its use in the scientific workflow with a hypothetical example project.
-In the tutorial below, we provide show how to use the Git VCS to curate your research materials by using text commands from the computers command line, and with a Graphical User Interface called R Studio.
+In the tutorial below, we show how to use the Git VCS by using text commands from the computer's command line.
+After introducing the fundamentals of Git's VCS functions from the command line, we also show how to use Git easily with a Graphical User Interface (GUI) implemented in the popular R Studio Integrated Development Environment (IDE).
 
@@
 It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (??
-Windows), Git is not a standalone program for navigating files and folders on a computer.
+Windows), Git[^git-name] is not a standalone program for navigating files and folders on a computer.
 Instead, it adds functionality to the OSs existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI) / integrated development environment (IDE)--that allow taking snapshots of the project and its files, and distributing the work across multiple computers and users.
@@
 [^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
+[^git-name]: The creator of Git, Linus Torvalds, named Git after himself as "the stupid content tracker" (@mcmillan_after_2005; <https://git-scm.com/docs/git.html>)
+[^simple]: By "simple", we only mean that the technical aspects of the project, related to curating materials related to it, are simple.
+We of course do not suggest that there might be anything theoretically or scientifically simple in such projects.
 

### 2017-06-15 manuscript/manuscript.Rmd
Edit # Version Control Systems

@@
 
-Version Control Systems (VCS; also known as Source Control systems) are computer programs designed for tracking changes to computer code, and sharing the code---and its history---with others.
-Although initially developed for writing computer code collaboratively, it is easy to recognize the benefits of VCS for scientific production and collaboration as well.
-For example, behavioral experiments are often written in a programming language, and can include multiple authors and versions.
+Version Control Systems (VCS; also known as Source Control Systems) are computer programs designed for tracking changes to computer code, and sharing the code with others.
+Although initially developed for writing computer code collaboratively[^vcs-collaborate], it is easy to recognize the benefits of VCS for scientific production and collaboration as well.
+For example, behavioral experiments are often written in a programming language, and can have multiple authors and versions.
 Keeping track of the versions of the program, and allowing many authors to contribute to it (without breaking the code) are problems that VCSs were specifically designed for.
-By extension, VCS easily adopts to other aspects of the research cycle, such as curating data across computers, laboratories, and experiments; even writing manuscripts is easier when it is recognized that manuscripts (usually) undergo multiple rounds of revision, and that multiple authors might want to keep track of all these different versions.
+By extension, VCS is easily adopted to other aspects of the research cycle, such as curating data across computers and laboratories, or writing manuscripts with multiple authors, versions, and sources of results.
 
-The core concepts of version control are that contributors to a project create small checkpoints of the changes they make to the source code (or manuscript text, data, etc.), and then submit those changes to the VCS.
-The VCS maintains a history of all these little checkpoints, and the exact state of the project at each of these checkpoints.
-
-<!-- Insert illustration of VCS about here. -->
-
-Some popular Version Control Systems are SVN, Mercurial, and Git.
-In this tutorial, we focus on the most widely used (in science) of these, called Git.
-Git is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.
-
-<!-- Examples of problems in psychology solved by VCS somewhere here. -->
-
+<!-- Examples of problems in psychology solved by VCS. -->
 \noindent
@@
   \begin{itemize}
-    \item Problem: Researchers usually want to try different ways of visualizing and modeling their data, and sometimes it is difficult to keep track of all the different versions.
+    \item Problem: Researchers want to try different ways of visualizing and modeling their data, and keep track of all the different versions they try.
       \begin{itemize}
@@
 
+The core concepts of version control are that contributors to a project create small checkpoints of the changes they make to the source code (or manuscript text, data, etc.), and then submit those changes to the VCS.
+The VCS maintains a history of all these little checkpoints, and the exact state of the project at each of these checkpoints.
+
+<!-- Insert illustration of VCS about here. -->
+
+Some popular Version Control Systems are SVN, Mercurial, and Git.
+In this tutorial, we focus on the most widely used (in science) of these, called Git.
+Git is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.
+
 ## The Git Version Control System
@@
 To begin using Git, users must first download and install the Git software on their computers.
-Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others), and is used by major software developers such as Microsoft, Google, and Facebook.
+Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others).
 It can be freely downloaded at <https://git-scm.com/>.
@@
 
+- Born open data [@rouder_what_2015]
+
 # Git Summary
@@
 We of course do not suggest that there might be anything theoretically or scientifically simple in such projects.
+[^vcs-collaborate]: The software we present below is used used by major software developers such as Microsoft, Google, and Facebook on code bases with hundreds of contributors.
 

### 2017-06-15 manuscript/manuscript.Rmd
Make proper tables.

@@
 output            : papaja::apa6_pdf
+header-includes:
+   - \usepackage{tcolorbox}
 ---
@@
 <!-- Examples of problems in psychology solved by VCS. -->
-\noindent
-\fbox{\begin{minipage}{\textwidth}
+\begin{tcolorbox}[colframe=green!50!black, colback=green!5, fonttitle=\small\bfseries, fontupper=\footnotesize, title=Box 1. Common problems and their VCS solutions]
     \begin{itemize}
-    \item Problem: Researchers want to try different ways of visualizing and modeling their data, and keep track of all the different versions they try.
+    \item How to try different ways of visualizing and modeling data while keeping track of the different versions---and the differences between them?
         \begin{itemize}
-        \item Solution: VCS allows researchers to save each version of the analysis into a history from where it can be easily recalled.
-This approach also eschews the need for creating multiple copies of the same analysis files with small tweaks, as illustrated \href{http://phdcomics.com/comics/archive.php?comicid=1531}{here}.
+        \item VCS saves each version of the analysis, allowing testing new features without losing previous versions or proliferating files in the project's directories.
+All the past versions can be directly compared with each other.
         \end{itemize}
@@
     \begin{itemize}
-    \item Problem:
+    \item How to work on the same code or manuscript files simultaneously with collaborators?
         \begin{itemize}
-        \item Solution:
+        \item VCS were built for distributed collaborative work: Collaborators work on their local copies and "push" and "pull" material to and from each other or a central "repository".
+VCS keeps precise track of who has done what, when, and (possibly) why.
         \end{itemize}
     \end{itemize}
-\end{minipage}}
+    \begin{itemize}
+    \item How to share my work in an organized manner with others?
+        \begin{itemize}
+        \item VCS requires committing to a common organization scheme with collaborators, making it easier to keep everyone "on the same page" with what goes where and how to contribute to specific parts of the project.
+Sharing the project with others is built into VCS, and can be facilitated with online services such as GitHub.
+        \end{itemize}
+    \end{itemize}
+\end{tcolorbox}
 
@@
 
-## Further resources for learning Git
-
-\noindent
-\fbox{\begin{minipage}{\textwidth}
+<!-- Box 2. Further resources for learning Git -->
+\begin{tcolorbox}[colframe=green!50!black, colback=green!5, fonttitle=\small\bfseries, fontupper=\footnotesize, title=Box 2. Further resources for learning Git]
 \begin{itemize}
-    \item \href{https://www.git-tower.com/blog/workflow-of-version-control}{Basic VCS workflow}, an infographic explaining how VCS works.
-    \item \href{https://services.github.com/resources/cheatsheets/}{GitHub's Git cheat sheet} is available in multiple languages and contains the most used Git commands.
-    \item \href{https://try.github.io/levels/1/challenges/1}{TryGit}, an interactive website for learning the basics of Git.
-    \item \href{http://r-pkgs.had.co.nz/git.html}{Git + GitHub} in an R programming context.
-    \item \href{https://git-scm.com/book/en/v2}{Pro Git book}, a complete manual of Git.
+    \item \href{https://www.git-tower.com/blog/workflow-of-version-control}{Basic VCS workflow}, an infographic explaining how VCS works (\url{https://www.git-tower.com/blog/workflow-of-version-control}).
+    \item \href{https://services.github.com/resources/cheatsheets/}{GitHub's Git cheat sheet} is available in multiple languages and contains the most used Git commands (\url{https://services.github.com/resources/cheatsheets/}).
+    \item \href{https://try.github.io}{TryGit}, an interactive website for learning the basics of Git (\url{https://try.github.io}).
+    \item \href{http://r-pkgs.had.co.nz/git.html}{Git + GitHub} in an R programming context (\url{http://r-pkgs.had.co.nz/git.html}).
+    \item \href{https://git-scm.com/book/en/v2}{Pro Git book}, a complete manual of Git (\url{https://git-scm.com/book/en/v2}).
 \end{itemize}
-\end{minipage}}
-
+\end{tcolorbox}
 \newpage

### 2017-06-15 manuscript/manuscript.Rmd
Add git-diagram (created with https://www.draw.io)

@@
 
-Version Control Systems (VCS; also known as Source Control Systems) are computer programs designed for tracking changes to computer code, and sharing the code with others.
-Although initially developed for writing computer code collaboratively[^vcs-collaborate], it is easy to recognize the benefits of VCS for scientific production and collaboration as well.
+Version Control Systems (VCS; also known as Source Control Systems) are computer programs designed for tracking changes to computer code, and collaborating on the code with others.
+Although initially developed for writing code collaboratively[^vcs-collaborate], it is easy to recognize that VCS can also be adopted for scientific production and collaboration.
 For example, behavioral experiments are often written in a programming language, and can have multiple authors and versions.
@@
 The VCS maintains a history of all these little checkpoints, and the exact state of the project at each of these checkpoints.
-
-<!-- Insert illustration of VCS about here. -->
+See Figure \@ref(fig:git-diagram) for a diagram of the typical VCS workflow.
 
@@
 These two operations are the backbone of using Git to store the state of the project whenever meaningful changes are made.
+
+```{r git-diagram, fig.cap = "A diagram illustrating the typical Git workflow. Verbs indicated in bold text are Git operations and explained in detail in the main text."}
+knitr::include_graphics("../data/git-diagram.png")
+```
 

### 2017-06-15 manuscript/manuscript.Rmd
Add diagram for collaborative workflow.

@@
 
+```{r git-diagram-collaborate, fig.cap = "A diagram illustrating the typical collaborative Git workflow with a remote repository (e.g.
+GitHub).
+Verbs indicated in bold text are Git operations and explained in detail in the main text.", out.width="6in"}
+knitr::include_graphics("../data/git-diagram-collaborate.png")
+```
+
 Setting up a centralized Git repository on a research team's private server is relatively straightforward, but because the details vary from team to team, here we illustrate the Centralized Workflow using GitHub.

### 2017-06-16 manuscript/manuscript.Rmd
Change Box 1 to talk about Git instead of VCS

@@
 
-<!-- Examples of problems in psychology solved by VCS. -->
-\begin{tcolorbox}[colframe=green!50!black, colback=green!5, fonttitle=\small\bfseries, fontupper=\footnotesize, title=Box 1. Common problems and their VCS solutions]
-    \begin{itemize}
-    \item How to try different ways of visualizing and modeling data while keeping track of the different versions---and the differences between them?
-        \begin{itemize}
-        \item VCS saves each version of the analysis, allowing testing new features without losing previous versions or proliferating files in the project's directories.
-All the past versions can be directly compared with each other.
-        \end{itemize}
-    \end{itemize}
-    \begin{itemize}
-    \item How to work on the same code or manuscript files simultaneously with collaborators?
-        \begin{itemize}
-        \item VCS were built for distributed collaborative work: Collaborators work on their local copies and "push" and "pull" material to and from each other or a central "repository".
-VCS keeps precise track of who has done what, when, and (possibly) why.
-        \end{itemize}
-    \end{itemize}
-    \begin{itemize}
-    \item How to share my work in an organized manner with others?
-        \begin{itemize}
-        \item VCS requires committing to a common organization scheme with collaborators, making it easier to keep everyone "on the same page" with what goes where and how to contribute to specific parts of the project.
-Sharing the project with others is built into VCS, and can be facilitated with online services such as GitHub.
-        \end{itemize}
-    \end{itemize}
-\end{tcolorbox}
-
 The core concepts of version control are that contributors to a project create small checkpoints of the changes they make to the source code (or manuscript text, data, etc.), and then submit those changes to the VCS.
@@
 Git is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.
+
+<!-- Examples of problems in psychology solved by Git. -->
+\begin{tcolorbox}[colframe=green!50!black, colback=green!5, fonttitle=\small\bfseries, fontupper=\footnotesize, title=Box 1. Common problems and their Git solutions]
+    \begin{itemize}
+    \item How to try different ways of visualizing and modeling data while keeping track of the different versions---and the differences between them?
+        \begin{itemize}
+        \item Git saves each version of the analysis, allowing testing new features without losing previous versions or proliferating files in the project's directories.
+All the past versions can be directly compared with each other.
+        \end{itemize}
+    \end{itemize}
+    \begin{itemize}
+    \item How to work on the same code or manuscript files simultaneously with collaborators?
+        \begin{itemize}
+        \item Git was designed for distributed collaborative work: Collaborators work on their local copies and "push" and "pull" material to and from each other or a central "repository".
+Git keeps precise track of who has done what, when, and (possibly) why.
+        \end{itemize}
+    \end{itemize}
+    \begin{itemize}
+    \item How to share my work in an organized manner with others?
+        \begin{itemize}
+        \item Git enforces a common organization scheme among collaborators, making it easier to keep everyone "on the same page" with what goes where, and how to contribute to specific parts of the project.
+Sharing the project with others is built into Git, and can be facilitated with online services such as GitHub.
+        \end{itemize}
+    \end{itemize}
+\end{tcolorbox}
 

### 2017-06-16 manuscript/manuscript.Rmd
Add basics of how to use command line

@@
 
-It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (??
-Windows), Git[^git-name] is not a standalone program for navigating files and folders on a computer.
-Instead, it adds functionality to the OSs existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI) / integrated development environment (IDE)--that allow taking snapshots of the project and its files, and distributing the work across multiple computers and users.
-To begin using Git, users must first download and install the Git software on their computers.
-Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others).
+Git[^git-name], unlike the operating system's (OS) default file viewer (Mac's Finder, Windows' File Explorer), is not a standalone program for navigating files and folders on a computer.
+Instead, Git adds functionality to the existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI)--that allow keeping track of files and their history, and distributing the files across multiple computers and users.
+Because Git is a standalone program and not usually included in standard OS installs, users must first download and install the Git software on their computers.
+Git is free and open source, works on Windows, Mac, and Linux OSs (among others).
 It can be freely downloaded at <https://git-scm.com/>.
-Before we explain the basics of using Git, we present brief installation instructions for Windows and Apple users.
+Before we explain the basics of using Git, we present brief installation instructions for Windows and Mac users.
 
@@
 
-Even if you already have Git installed (some computers do) it is a good idea to install the latest version[^git-version], which can be downloaded as a standalone program from <https://git-scm.com/download>.
-Here, we provide more detailed instructions on installing Git for OS X and Windows, but Linux (and other OS) users can find instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>).
-After installation instructions, we detail how to configure the Git program so that it is ready to be used.
+Even if you already have Git installed (some computers do) it is good practice to install the latest version[^git-version], which can be downloaded as a standalone program from <https://git-scm.com/download>.
+Here, we provide more detailed instructions on installing Git for Mac and Windows OSs, but Linux (and other) users can find instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>).
+After installation instructions, we detail how to configure the Git program so that you can start using it.
 
-### OS X
+### Installing Git on Macs
 
-Many OS X computers already have Git installed, especially those operating OS X 10.9 or higher, but it is good practice to install the latest version.
+Some Mac computers already have Git installed, especially those operating OS X 10.9 or higher, but it is good practice to install the latest version.
 The easiest way to update Git to the latest version is to download the installer from <http://git-scm.com/download/mac>, and install it like any other program.
-Once Git is installed, it can be used from the command line (the OS X command line is available through the Terminal app, which is included by default with the OS X install), or through various GUIs.
-Although it might at first appear intimidating because of the archaic-looking interface, we encourage using Git from the command line.
 
-### Windows
+### Installing Git on Windows
 
@@
 
-To operate properly, Git needs to be able to identify its user.
-You can set these by entering a few basic commands in the command line.
-To verify the current user information, type `git config --global user.name`, and hit return.
+Once Git is installed, it can be used from the command line or through a GUI--there are various GUIs for using Git, and below we will explain how to use the R Studio Git GUI.
+First, it is important to learn a few basic Git commands from the OSs command line.
+The command line is a text-based interface for interacting with your computer, and its functionality greatly extends that of the standard way of interacting with the computer by clicking and pointing with a mouse.
+
+To access the command line interface, you need to use a command line "shell".
+Mac users can open the built-in app *Terminal*, and Windows users can use one of many alternatives, such as Microsoft PowerShell (<https://msdn.microsoft.com/powershell>).
+After opening the shell app, you can type in commands to access the interface.
+For example typing `ls` (hit Return to execute) lists the files in the current folder.
+
+The first step in using Git is making Git aware of who is using the computer.
+You need to set the user name and email by entering a few basic commands in the command line.
+First, to show the current user information, type `git config --global user.name`, and hit return.
 This should not return anything, unless a previous user of the computer has set the global Git user name.
 
+Each Git command starts with the word `git`, then a command (such as `config`), and then arguments to the command, such as `--global` (for global configuration) followed by the `user.name` variable name to show the global user name.
+For detailed instructions on how to use Git commands, you can type `git --help`.
+For help on how to use the `git config` command, type `git config --help`.
+
 To ensure that Git knows who you are, type `git config --global user.name "User Name"` (where User Name is your name) in the command line, and hit return.
-If you now re-run the first command, the command line will return the name you entered as `"User Name"`.
-The username can be anything you'd like, but it is probably a good idea to use your real name so that potential collaborators know who you are.
-The second piece of information is your email, which can be entered by `git config --global user.email "email@address.com"` (where email@address.com is your email address).
+This command maps `"User Name"` to Git's global `user.name` variable.
+If you now re-run the first command (`git config --global user.name`), the command line will return the name you entered as `"User Name"`.
+The user name can be anything you'd like, but it is probably a good idea to use your real name so that potential collaborators know who you are.
+The second piece of information is your email address, which is entered by `git config --global user.email "email@address.com"` (where `"email@address.com"` is your email address).
 You can verify that the correct email address was saved by typing `git config --global user.email` in the Terminal, and hit return.
-Once this information is entered, Git will know who you are, and is thus able to track who is doing what within a project, which is especially helpful when you are collaborating with other people, or when you are working on multiple computers.
+
+Once this information is entered, Git will know who you are, and is able to track who is doing what and when within a project, which is especially helpful when you are collaborating with other people, or when you are working on multiple computers.
+
+For users who have not previously used text-based command line interfaces, typing out these commands may seem intimidating, but to get started you only need two configuration commands; the basic use of Git requires only 4 commands.
+If using the text-based command line seems a daunting task, Codecademy (<https://www.codecademy.com/learn/learn-the-command-line>) has a free interactive online tutorial, and MIT has an online game to teach using the command line (<http://web.mit.edu/mprat/Public/web/Terminus/Web/main.html>).
 

### 2017-06-16 manuscript/manuscript.Rmd
Clarified writing up to # Using Git

@@
 Reproducibility can be defined as follows: "A research project is computationally reproducible if a second investigator (including you in the future) can recreate the final reported results of the project, including key quantitative findings, tables, and figures, given only a set of files and written instructions." [@kitzes_practice_2017].
-In the context of experimental Psychology, for example, a "project" could be an experiment or a set of experiments investigating a theory or hypothesis, "reported results" could be figures or statistical analyses reported in a conference presentation or a manuscript.
+In the context of experimental Psychology, for example, a "project" could be an experiment or a set of experiments investigating a hypothesis, "reported results" could be figures or statistical analyses in a conference presentation or a manuscript.
 
@@
 To put it more generally, replicability is a broad conceptual issue concerning the epistemology of scientific claims.
-Reproducibility, on the other hand, is a technical requirement for allowing others (including researchers themselves) to assess the reliability of a specific set of empirical or computational results.
+Reproducibility, on the other hand, is a technical requirement for allowing others (including researchers themselves in the future) to assess the reliability of a specific set of reported empirical or computational results.
 
-Further, although they are often erroneously conlated [@open_science_collaboration_estimating_2015; @peng_reproducible_2011] reproducibility and replicability are independent of one another.
+Further, although they are sometimes erroneously conlated [@open_science_collaboration_estimating_2015; @peng_reproducible_2011] reproducibility and replicability are independent of one another.
 Although the distinction between the two may be somewhat blurred in purely computational areas (e.g.
@@
 Given this broad definition of reproducibility, it is easy to recognize that many projects fall between the extremes of completely replicable and not replicable at all.
-A completely reproducible project would provide consumers the entire raw data set, and detailed instructions to analyze it (usually as unambigous computer code).
+A completely reproducible project would provide consumers the entire raw data set, and detailed, unambiguous instructions to analyze it (usually as computer code).
 This degree of reproducibility is very rare in Psychology [@wicherts_poor_2006; @vanpaemel_are_2015], although data sharing has become more common in recent years in journals that award authors who do so with special badges [@kidwell_badges_2016].
@@
 Vanpaemel et al.
-highlight some reasons provided by authors for not sharing their data:
+[-@vanpaemel_are_2015] highlight some reasons provided by authors for not sharing their data:
 
@@
 This state of affairs is not surprising because researchers are trained to do research, but receive little or no training in formal methods or accepted gold standards for curating their research assets.
-For simple[^simple] projects, such as writing a solo-author manuscript, this has not been a problem.
-However, with increasing complexity of research designs, number of collaborators, research sites, and technical skills required in modern scientific workflows, not knowing how to keep track of materials can lead to serious problems, as detailed above.
+For simple[^simple] projects, such as writing a solo-author manuscript, this may not be a problem.
+However, with increasing complexity of the type of data collected by behavioral scientists, and the increased number of collaborators, research sites, and technical skills required in modern scientific workflows, not knowing how to curate research assets can lead to wasted time or worse problems, as detailed above.
 
@@
 
-Version Control Systems (VCS; also known as Source Control Systems) are computer programs designed for tracking changes to computer code, and collaborating on the code with others.
+Version Control Systems are computer programs designed for tracking changes to computer code, and collaborating on the code with others.
 Although initially developed for writing code collaboratively[^vcs-collaborate], it is easy to recognize that VCS can also be adopted for scientific production and collaboration.
-For example, behavioral experiments are often written in a programming language, and can have multiple authors and versions.
-Keeping track of the versions of the program, and allowing many authors to contribute to it (without breaking the code) are problems that VCSs were specifically designed for.
-By extension, VCS is easily adopted to other aspects of the research cycle, such as curating data across computers and laboratories, or writing manuscripts with multiple authors, versions, and sources of results.
+For example, behavioral experiments are often created with a programming language, and can have multiple authors and versions.
+Keeping track of the versions of the program and changes to the code, and allowing many authors to contribute to it (without breaking the experimental program) are problems that VCSs were specifically designed for.
+VCS is also easily adopted to other aspects of the research cycle, such as curating data across computers and laboratories, or writing manuscripts with multiple authors, versions, and sources of results.
 
-The core concepts of version control are that contributors to a project create small checkpoints of the changes they make to the source code (or manuscript text, data, etc.), and then submit those changes to the VCS.
-The VCS maintains a history of all these little checkpoints, and the exact state of the project at each of these checkpoints.
+The core concepts of version control are that contributors to a project create small checkpoints of the changes they make to the source code---analogous to saving an intermediate version of a file on the computer's hard disk---and then submit those changes to the VCS.
+The VCS maintains a history of changes to the code between these little checkpoints, and therefore allows coming back to any earlier version by browsing the history.
 See Figure \@ref(fig:git-diagram) for a diagram of the typical VCS workflow.
@@
 Some popular Version Control Systems are SVN, Mercurial, and Git.
-In this tutorial, we focus on the most widely used (in science) of these, called Git.
-Git is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.
+In this tutorial, we focus on Git: Git is already incresing in popularity in the scientific community, and is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.
 
 <!-- Examples of problems in psychology solved by Git. -->
-\begin{tcolorbox}[colframe=green!50!black, colback=green!5, fonttitle=\small\bfseries, fontupper=\footnotesize, title=Box 1. Common problems and their Git solutions]
+\begin{tcolorbox}[colframe=green!50!black, colback=green!5, fonttitle=\small\bfseries, fontupper=\footnotesize, title=Box 1. How Git facilitates the scientific workflow.]
     \begin{itemize}
@@
 
-Git[^git-name], unlike the operating system's (OS) default file viewer (Mac's Finder, Windows' File Explorer), is not a standalone program for navigating files and folders on a computer.
-Instead, Git adds functionality to the existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI)--that allow keeping track of files and their history, and distributing the files across multiple computers and users.
-Because Git is a standalone program and not usually included in standard OS installs, users must first download and install the Git software on their computers.
-Git is free and open source, works on Windows, Mac, and Linux OSs (among others).
-It can be freely downloaded at <https://git-scm.com/>.
-Before we explain the basics of using Git, we present brief installation instructions for Windows and Mac users.
+Git[^git-name], unlike the operating system's (OS) default file viewer (Mac's Finder, Windows' File Explorer), is not a point-and-click program for navigating files and folders on a computer.
+Instead, Git adds functionality to the existing file system, by making available a specific set of commands---either executed from the command line, or through a graphical user interface (GUI)---which allow keeping track of files and their history, and distributing the files across multiple computers and users.
+Because Git is a standalone program not usually included in standard OS installs, users must first download and install the Git software on their computers.
+Git is free and open source, works on Windows, Mac, and Linux OSs (among others), and can be freely downloaded at <https://git-scm.com/>.
 
@@
 Even if you already have Git installed (some computers do) it is good practice to install the latest version[^git-version], which can be downloaded as a standalone program from <https://git-scm.com/download>.
-Here, we provide more detailed instructions on installing Git for Mac and Windows OSs, but Linux (and other) users can find instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>).
-After installation instructions, we detail how to configure the Git program so that you can start using it.
-
-### Installing Git on Macs
-
-Some Mac computers already have Git installed, especially those operating OS X 10.9 or higher, but it is good practice to install the latest version.
-The easiest way to update Git to the latest version is to download the installer from <http://git-scm.com/download/mac>, and install it like any other program.
-
-### Installing Git on Windows
-
-Windows users can download the Git software installer from <http://git-scm.com/download/win>. [TODO]
+For Mac users, the easiest way to install or update Git to the latest version is to download the installer from <http://git-scm.com/download/mac>, and install it like any other program.
+Windows users can download the Git software installer from <http://git-scm.com/download/win>.
+Linux (and other) users can download Git and find further install instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>).
 
@@
 
-Once Git is installed, it can be used from the command line or through a GUI--there are various GUIs for using Git, and below we will explain how to use the R Studio Git GUI.
+Once you have installed Git, you can be use it from the command line or through a GUI--there are various GUIs for using Git, and below we will explain how to use the R Studio Git GUI.
 First, it is important to learn a few basic Git commands from the OSs command line.
@@
 
-To access the command line interface, you need to use a command line "shell".
-Mac users can open the built-in app *Terminal*, and Windows users can use one of many alternatives, such as Microsoft PowerShell (<https://msdn.microsoft.com/powershell>).
+To access the command line interface, you need to use a command line "shell" application.
+Mac users can open the built-in app *Terminal*, and Windows users can download and use one of many alternatives, such as Microsoft PowerShell (<https://msdn.microsoft.com/powershell>).
 After opening the shell app, you can type in commands to access the interface.
 For example typing `ls` (hit Return to execute) lists the files in the current folder.
+Some users may find using a text-based command line interface unfamiliar, but to get started with Git, there are only two required configuration commands which you need to run once, and the basic functionality requires using only four commands[^cli-links].
+However, users can also read through the first part of the tutorial without executing the commands and simply focus on the workflow; we show how to use Git with a GUI in the second part of the tutorial.
 
 The first step in using Git is making Git aware of who is using the computer.
-You need to set the user name and email by entering a few basic commands in the command line.
+You need to set the user name and email address by entering a few basic commands in the command line.
 First, to show the current user information, type `git config --global user.name`, and hit return.
@@
 Once this information is entered, Git will know who you are, and is able to track who is doing what and when within a project, which is especially helpful when you are collaborating with other people, or when you are working on multiple computers.
-
-For users who have not previously used text-based command line interfaces, typing out these commands may seem intimidating, but to get started you only need two configuration commands; the basic use of Git requires only 4 commands.
-If using the text-based command line seems a daunting task, Codecademy (<https://www.codecademy.com/learn/learn-the-command-line>) has a free interactive online tutorial, and MIT has an online game to teach using the command line (<http://web.mit.edu/mprat/Public/web/Terminus/Web/main.html>).
 
@@
 [^vcs-collaborate]: The software we present below is used used by major software developers such as Microsoft, Google, and Facebook on code bases with hundreds of contributors.
+[^cli-links]: If using a text-based command line seems challenging, Codecademy (<https://www.codecademy.com/learn/learn-the-command-line>) has a free interactive online tutorial, and MIT offers a free online game to teach using the command line (<http://web.mit.edu/mprat/Public/web/Terminus/Web/main.html>).
 

### 2017-06-16 manuscript/manuscript.Rmd
Clarify writing of part I of tutorial

@@
 In the tutorial below, we show how to use the Git VCS by using text commands from the computer's command line.
-After introducing the fundamentals of Git's VCS functions from the command line, we also show how to use Git easily with a Graphical User Interface (GUI) implemented in the popular R Studio Integrated Development Environment (IDE).
+After introducing the fundamentals of Git's VCS functions from the command line, we also show how to use Git easily with a Graphical User Interface (GUI) implemented in the popular R Studio Integrated Development Environment (IDE; @rstudio_team_rstudio:_2016).
 
@@
 
-The first operating principle of Git is that your work is organized into independent projects, which Git calls *repositories*.
-At its core, a repository is a folder on your computer, which is version controlled by Git (you can tell if a folder is "monitored" by git by checking if the folder, or any of its parent folders, contains a hidden `.git` directory).
-Everything that happens inside a repository is tracked by Git, but the user has full control of what is tracked (everything, by default) and when.
-Because the user(s) has full control of what and when is tracked, there is a small set of operations the users need to execute every time they wish to log something in Git.
+The first operating principle of Git is that your work is organized into independent projects, which Git calls *repositories*[^git-link-submodules].
+A repository is a folder on your computer which is version controlled by Git (you can tell if a folder is "monitored" by git by checking if the folder, or any of its parent folders, contains a hidden `.git` directory).
+Everything that happens inside a repository is tracked by Git, but you have full control of what is committed to Git's history and when.
+Because you have full control of what and when is committed to history, there is a small set of operations you need to know[^git-cheat-sheet].
 
-Briefly, when you work in a Git repository, Git stores the state of each file that it monitors, and when any of these files change, Git indicates that they differ from the previously logged state.
-If the user then is happy with the current changes, she can **add** the changed files to Git's "staging area".
-If the user then is certain that the files added to the staging area are in good shape, they can **commit** the changes.
+Briefly, when you work in a Git repository (make changes to files within it), Git monitors the state of each file, and when they change Git knows that they differ from the previously logged state.
+If you are happy with the current changes, you **add** the changed files to Git's "staging area".
+If you then are certain that the changes in the staging area are desirable, you **commit** the changes.
 These two operations are the backbone of using Git to store the state of the project whenever meaningful changes are made.
+Importantly, each commit in the repository's history contains information to recover the full state of the project at that point in time.
+Users can always go back to an earlier version by **checking out** a previous state from Git's history.
+This core of the Git workflow is illustrate in Figure \@ref(fig:git-diagram).
 
@@
 
-Most importantly, each state in Git's log contains the full state of the files at that point in time.
-Users can always go back to an earlier version by "checking out" a previous state of Git's log.
-That's why these programs are called Version Control Systems.
-
-To establish the operating principles of Git in practice, we now turn to a practical example using a hypothetical project that, we feel, reflects the components of a typical project in experimental psychology.
-To most clearly show the use of Git, we begin with an empty project with no components.
+To understand the Git workflow in practice, we now turn to a practical example using a hypothetical project.
+Git can be added to a project at any stage of the project's life cycle, but to most clearly show its use, we begin with an empty project with nothing in it.
 
@@
 
-Implementing reproducibility into the scientific workflow is less time-consuming if it is planned from the onset of the project, rather than attempting to add reproducibility to the project after it has been completed.
-It is therefore important to organize the project keeping a few key goals in mind (here, we broadly follow established guidelines such as Project TIER recommendations): The files and folders should have easy to understand names (avoid idiosynchratic naming schemes), and the names should indicate the purposes of the files and folders.
+Implementing reproducibility into the scientific workflow is less time-consuming and effortful if it is planned from the onset of the project, rather than added to the project after all the work has been completed.
+It is therefore important to organize the project keeping a few key goals in mind (here, we follow guidelines such as the Project TIER recommendations[^tier-link]): The files and folders should have easy to understand names (avoid idiosynchratic naming schemes), and the names should indicate the purposes of the files and folders.
 
@@
 All the materials related to this project will be placed in subfolders of the home directory, but for now `git-example` is just an empty folder waiting to be filled with data, code, and documents.
+To make the example concrete, you can follow the tutorial by typing out the following commands on your own computer.
 
@@
 
-Next, you need to navigate to the folder and initialize it as a Git repository.
-Here, the `git-example` project is in the users `Documents` folder, and we can use `cd ..` in the OS X terminal to navigate up in the folder hierarchy, and `cd <folder>` to navigate into `<folder>`.
-So assuming that the Terminal opens up in `Documents`, we navigate to `git-example` with (to execute code in Terminal, enter the text into the command line and hit Return):
+Next, you need to create a new folder for the Git repository.
+First, choose an appropriate folder on your computer (such as `User/Documents/`) where you'd like to create the project.
+You can either use the system's file navigator (Finder / File Explorer) to create this folder, or the following commands in the command line: Navigate to the desired folder by using `cd Documents` to move into the `Documents` folder (assuming it exists in the folder where you currently are), and `cd ..` to move out of the folder (to its containing folder.) Once you are in the folder where you want to create the project, type `mkdir git-example` to make the `git-example` directory, then `cd git-example` to move into it.
+Once you are in the project's home folder (you can verify where you are by typing `pwd`), you want to turn the folder into a project by initializing Git:
 
 ```bash
-cd git-example
+> git init
 ```
 
-Then, to turn this folder into a Git repository, execute
-
-```bash
-git init
-```
-
-This command initializes the folder as a Git repository, and the only change so far has been the addition of a hidden `.git` folder inside `git-example` (and a `.gitattributes` file; but users can ignore these hidden files and folders).
+This command initializes the folder as a Git repository, and the only change so far has been the addition of a hidden `.git` folder inside `git-example` (and a `.gitattributes` file.
+Users can ignore these hidden files and folders, however they can be shown by typing `ls -la`.) Now that the folder is initialized as a Git repository, Git monitors any changes within it, and allows you to add and commit these changes (Figure \@ref(fig:git-diagram).)
 
@@
 
-Every project, and therefore every repository, should contain a brief not that explains what the project is about and who to contact about it.
-This note is usually called a readme, and therefore our first contribution to this project will be a README file.
+Every project (repository) should contain a brief note explaining what the project is about and who to contact.
+This note is usually called a readme, and therefore our first contribution to this project will be a README file (this file is so important that it has become standard practice to write it in capital letters).
 The README file should be a plain text file (i.e.
 not created with Microsoft Word) that can be read with a simple text editor.
-This file can now be seen in the `git-example` folder by using the system file viewer.
+You can create this file with any text editor (Macs have TextEdit app installed by default, which allows creating plain text files).
+We created the file with some text in it, and it can now be seen in the `git-example` folder, either by using the system file viewer or `ls` in the command line.
 Because we added this file to a Git repository, Git is also aware of it.
@@
 ```bash
-git status
-```
+> git status
 
-Which in this case would return
-
-```bash
-Matti [~/Documents/git-example]$ git status
 On branch master
@@
 nothing added to commit but untracked files present (use "git add" to track)
-Matti [~/Documents/git-example]$ 
 ```
@@
 The relevant output returned from executing this command is the "Untracked files:" part.
-There, Git tells the user that there is an untracked file (README) in the repository.
-To keep track of the status of this file, we **add** it to Git by using the command `git add` followed by either a `.` (for adding all untracked files) or `README` for only tracking the `README` file.
+There, Git tells that there is an untracked file (README) in the repository.
+To start tracking changes in this file, we **add** it to Git by using the command `git add` followed by either a `.` (for adding all untracked files) or `README` for only tracking the `README` file.
 
 ```bash
-git add README
+> git add README
 ```
 
-We've now added this file to the staging area, and if we are happy with changes to the file's status (it has been created), we can **commit** the file to Git's versioning system.
-Commits are the most important Git operation: They signify any meaningful change to the repository, and can be later browsed and compared to one another.
+We've now added this file to the staging area (Figure \@ref(fig:git-diagram)), and if we are happy with changes to the file's status, we can **commit** the file to Git's history.
+Commits are essential Git operations: They signify meaningful changes to the repository, and can be later browsed and compared to one another.
 As such, it is helpful to attach a small message to each commit, describing why that commit was made.
@@
 ```bash
-git commit -m "Add README file."
+> git commit -m "Add README file."
 ```
 
-The quoted text after the `-m` flag is the "commit message".
+The quoted text after the `-m` argument is the *commit message*.
 Entering this command to the command line returns a brief description of the commit, such as how many files changed, and how many characters inside those files were inserted and deleted.
@@
 
-Most importantly, the `git-example` project now keeps track of all and any changes to README.
-To illustrate, we now add some text to README to describe the project (title, what the project is about, who are involved, who to contact), save the file, and then ask for Git's status with `git status` on the command line:
+The `git-example` project (or rather, Git) now keeps track of all and any changes to README.
+To illustrate, you can change the text in the README file with a text editor, save the file, and then ask for Git's status with `git status` on the command line:
 
 ```bash
-Matti [~/Documents/git-example]$ git status
+> git status
+
 On branch master
@@
   (use "git checkout -- <file>..." to discard changes in working directory)
-
 	modified:   README
-
 no changes added to commit (use "git add" and/or "git commit -a")
@@
 
-Git can now tell that the README file has changed, and we can repeat the add and commit steps to permanently record the current state of the project to Git's log:
+Git can tell that the README file has changed.
+It is often useful to know exactly *how* a file has changed, before committing it.
+To view **diff**erences to a file not yet committed, use `git diff <file>`.
+It shows changes within `<file>`, line by line, highlighting removals with red and added lines with green.
+Once you are happy with the changes, you can repeat the add and commit steps from above to permanently record the current state of the project to Git's log (below we use the `.` shortcut for adding all changes):
 
 ```bash
-git add .  # We use the '.' shortcut 
-git commit -m "Populate README with project description."
+> git add .
+> git commit -m "Populate README with project description."
 ```
@@
 
-The real importance of these somewhat abstract steps becomes apparent when we consider the Git **log**. We can call
+The real importance of these somewhat abstract steps becomes apparent when we consider the Git **log**. To reveal the commit log of your repository, call:
 
 ```bash
-git log
+> git log
 ```
 
-To reveal the commit log of this repository.
-The main things to notice in the output of this command are that each commit is associated with a unique hash code (long alphanumeric string), which we can use to call for further information (see below); an author (this is where the earlier setup is apparent); a date and time; and the short commit message.
-Until now, the `git-example` log looks like this:
+The main outputs of this command show that each commit is identified with a unique hash code (long alphanumeric string), which we can use to call for further information (see below); an author; a date and time; and the short commit message.
+Executing `git log` on our example project at this stage returns this:
 
 ```bash
-Matti [~/Documents/git-example]$ git log
 commit 60cbe5c9b4a78e500314f791080381030577a035
@@
 Date:   Tue Jun 13 17:20:27 2017 -0400
-
     Populate README with project description.
@@
 Date:   Tue Jun 13 17:14:14 2017 -0400
-
     Add README file.
@@
 
-To see what exactly changed in the last commit (the log has latest commits at the top), we can call `git show` with the commit's hash (only relevant parts of output shown below):
+To see what exactly changed in the last commit (the log has latest commits at the top), you can call `git show` with the commit's hash (only relevant parts of output shown below):
 
 ```bash
-Matti [~/Documents/git-example]$ git show 60cbe5c9b4a78e500314f791080381030577a035
+> git show 60cbe5c9b4a78e500314f791080381030577a035
 commit 60cbe5c9b4a78e500314f791080381030577a035
@@
 Date:   Tue Jun 13 17:20:27 2017 -0400
-
     Populate README with project description.
-
 diff --git a/README b/README
@@
 
-This output can be investigated for a detailed log of all changes created by that commit.
-From top, it lists the author of the commit, the commit message, and then the commit's "**diff**" (i.e.
+This output is a detailed log of all changes in that commit.
+From top, it lists the author, the message, and then the commit's "**diff**" (i.e.
 what differs in the new version vs the old version of the file.) The current diff shows that 1 file received eight additional lines of text (the part wrapped in @ symbols), and then the additions themselves (the lines prepended with +s).
 
-Although we now understand the fundamentals of using Git to track the states of (and therefore changes to) a repository, this contrived and overly simplistic example doesn't allow full appreciation of the benefits of using Git for version control.
+Although you have now seen the fundamentals of using Git to track the states of (and therefore changes to) a repository, this contrived and overly simplistic example doesn't allow full appreciation of the benefits of using Git for version control.
 To better illustrate Git's functioning, we now fast-forward in the hypothetical example project to a stage in the project where more files and materials have been created.
 
-## (Slightly more) advanced Git
+## (Slightly More) Advanced Git
 
-The state of the project now looks like this:
+For example, after working for a while on the project, you could have added two files inside well-organized subfolders of the project. The project would like this when viewed with a file explorer:
 
@@
 git-example/
-    admin/
-        ethics-info.pdf
-    experiment-1/
-        analysis/
-            plan_n.R
-    manuscript/
-    README
+|-- admin/
+    |-- ethics-info.pdf
+|-- experiment-1/
+    |-- analysis/
+        |-- plan_n.R
+|-- manuscript/
+|-- README
 ```
 
-Running `git status` now tells that there are two new files (empty folders are ignored).
-That is, since the previous commit, two files have been added to the project; one a pdf with some administrative information (`ethics-info.pdf`), the other one is an R script for a pre-planned power analysis (`plan_n.R`).
-We would certainly like to track any changes to the power analysis script, but the ethics information file isn't really something that we need to keep track of---we can't make changes to it anyway.
+Running `git status` now tells that there are two new files (possible empty folders are ignored).
+That is, since the previous commit, two files have been added to the project; one a pdf with some administrative information (`ethics-info.pdf`), the other one is an R [@r_core_team_r:_2017] script for a pre-planned power analysis (`plan_n.R`).
+You would probably like to track any changes to the power analysis script, but the ethics information file isn't really something that you need to keep track of---at least for this example.
 It would be laborious to constantly keep ignoring it when committing changes to Git (especially if we'd like to commit multiple files simultaneously with `git add .`) Git has an elegant solution to specifying which files to keep track of.
-Because by default all files are tracked, Git uses a special file for instructing which files are to be *ignored* from tracking.
+Because by default all files are monitored, Git uses a special file for instructing which files are to be *ignored*.
 
@@
 
-To make Git ignore files, we simply add a plain text file called `.gitignore` to the home folder of the repository.
+To make Git ignore files, you simply add a plain text file called `.gitignore` to the home folder of the repository.
+You can use any text editor to create this file.
 Each row of this file should specify a file or a folder (or a regular expression) that Git should ignore.
-In our example we want to make Git ignore the `admin/` folder entirely, and any file with the .pdf extension inside `manuscript/.
-The example .gitignore file looks like this (lines beginning with `#` are comments):
+In the current example you could make Git ignore the `admin/` folder entirely, and any file with the .pdf extension inside `manuscript/`.
+The example .gitignore file would look like this (lines beginning with `#` are comments):
 
@@
 manuscript/*.pdf
+
+# Ignore a specific file in a specific folder
+analysis/sandbox.R
 ```
@@
 
-Because there are now two untracked files, and we would like to keep a clean history of what has happened in the project, we create two separate commits: One for the .gitignore file, and one for the power analysis file.
+Because there are now two untracked files that are not in .gitignore (`.gitignore` and `plan_n.R`), and you usually should aim to maintain a clean history for the project, you can create two separate commits: One for the Git ignore file, and one for the power analysis file.
 
@@
 
-After this last commit, if we ever want to know in the future what our planned sample size was, we can come back to this commit with `git log` or `git show` and see what was inside the newly created power analysis file.
-For instance, if new information suggests that we should change assumed effect size in the power analysis, we can simply edit and save the file, then add and commit the changes to Git with a helpful message that logs this important event in the Git memory of the repository.
+After this last commit, you can at any time come back to this commit with `git log` or `git show` and see what was inside the newly created power analysis file when it was first created.
+For instance, if new information suggests that you should change the assumed effect size in the power analysis, you can simply edit and save the file, then add and commit the changes to Git with a helpful message that logs this important event in Git's history.
 
-This possibility of "rewinding history" is especially useful for files that might go under multiple revisions, or if we might be interested when and in what order the files were created.
+This possibility of "rewinding history" is especially useful for files that might go under multiple revisions (manuscripts, analysis files), or if you are interested when and in what order the files were created---and who created or changed them.
 One might consider committing a power analysis as a small personal pre-registration of a part of the research plan.
@@
 [^cli-links]: If using a text-based command line seems challenging, Codecademy (<https://www.codecademy.com/learn/learn-the-command-line>) has a free interactive online tutorial, and MIT offers a free online game to teach using the command line (<http://web.mit.edu/mprat/Public/web/Terminus/Web/main.html>).
+[^git-link-submodules]: For advanced users, Git *submodules* allow linking projects to each other, or organizing more complex projects into projects and their sub-projects (<https://git-scm.com/book/en/v2/Git-Tools-Submodules>).
+[^git-cheat-sheet]: It helps to have a Git command cheat sheet (<https://services.github.com/resources/cheatsheets/>) printed and taped on your wall, but it contains many more commands than are needed for the basic use of Git.
+[^tier-link]: <http://www.projecttier.org/tier-protocol/specifications/>
 

### 2017-06-16 manuscript/manuscript.Rmd
Add  ### Try new feature and finish  # Collaborating

@@
 
+### Try a new feature
+
+We often find that making some changes to a project didn't have the desired effect: The manuscript ended weaker or the analysis didn't work anymore.
+Git allows great flexibility in trying new features, then undoing the changes[^git-branches].
+Starting with an empty staging area, you could start modifying a file (e.g.
+`plan_n.R`), and after a while find out that the changes were not for good.
+At this point it is common to press "Undo" in the text editor, but if the file has been saved multiple times or multiple files have been changed, it is difficult to get to the starting point by simply using the "Undo" button.
+Instead, with Git you can **checkout** the file's previous version from history.
+To undo all changes to `plan_n.R` (since the last commit), run
+
+```bash
+git checkout experiment-1/analysis/plan_n.R
+```
+
+Notice that you have to write the full path of the file (relative to the project's root) so Git knows precisely which file you want to checkout from history.
+With these example operations, we have discussed the main Git operations as outlined in Figure \@ref(fig:git-diagram): Make changes to files, **add** to staging area, **commit** to history; **checkout** from history to undo changes.
+
 # Collaborating
 
-The true advantages of using a VCS such as Git become apparent when we consider projects with more than one contributor.
+The true advantages of using Git become apparent when we consider projects with more than one contributor.
 For example, consider a project where data is collected at multiple sites, and these files are then saved onto a centralized server, or shared through a service that automatically merges files from multiple sources (such as the popular Dropbox service).
@@
 
-Or consider a data analysis where two or more people work simultaneously on some complicated analysis.
-If user A and B have the same file open on multiple computers, after a while when user B saves the file, user A's version of the file will be overwritten.
+Alternatively, consider a data analysis where two or more people work simultaneously on some complicated analysis.
+If user A and B are making changes to the same file and user B saves the file, user A's version of the file will be overwritten.
 Disaster!
 
-Git and other VCSs were specifically designed to allow (and facilitate) multi-site collaboration on extremely complex projects, as for example required by leading software companies such as Apple and Microsoft.
-Next, we'll illustrate how collaborating on a Git project works in practice.
+Git and other VCSs, on the other hand, were specifically designed to allow (and facilitate) multi-site collaboration on arbitrarily complex projects: Microsoft Windows is developed on a Git platform.
+Git is especially helpful in scientific collaboration, a topic to which we turn next.
 
 There are many ways in which a team could collaborate on a Git project (e.g.
-<https://www.atlassian.com/git/tutorials/comparing-workflows>), and here we focus on a common one, called the "Centralized Workflow".
+<https://www.atlassian.com/git/tutorials/comparing-workflows>); here we focus on a common one, called the "Centralized Workflow" where one virtual copy of the project is considered the central node, and all contributors interact with this central node, instead of directly with each other.
 
@@
 
-In this workflow, a central repository hosted on a research team's server, or an online service like GitHub (<https://github.com/>), is used to pull and push (see Table \@ref(tab:git-commands)) information to and from local repositories.
-First, one user creates this central repository, then other users clone a local copy (see Table \@ref(tab:git-commands)).
-All changes are first created, added, and committed locally as detailed above, and only then **pushed** to the central repository.
-To get changes that other users have pushed to the central repository, a user **pulls** changes from it.
+In this workflow, a central project hosted on a research team's server, or an online service like GitHub (<https://github.com/>), is used to send and receive information from and to local projects.
+Because the word "project" can now refer to virtual copies of projects, they are instead commonly referred to as **repositories**.
+First, one user creates this central repository, then other users **clone** a local copy of it (see Table \@ref(tab:git-commands)).
+Contributors, including the one who created the central repository, then work on their local projects as detailed above, and after committing **push** their changes to the central repository.
+To get changes that other users have pushed to the central repository, contributors **pull** changes from it.
 
@@
 GitHub is one of the 100 most popular website worldwide, and hosts over 60 million software projects with a total of over 20 million users (<https://github.com/about>).
-We chose GitHub for collaboration because it is already the de-facto standard in VCS collaboration among scientists, it offers free private (see below) repositories for students, and has many attractive features, some of which we will detail below.
+We chose GitHub for collaboration because it is already the de-facto standard in VCS collaboration among scientists, it offers free private (see below) repositories for students, and repositories from GitHub can easily be connected to projects hosted on the Open Science Framework (<https://osf.io>).
 
@@
 
-Although we are creating a new repository here, the goal is to create an online "remote" for the running example within this tutorial, and therefore the specific steps are slightly different than what they would be if a brand new repository was needed.
-
-For the current example, after clicking "New repository", we simply enter a name, which can be anything, but for consistency we call the GitHub repository `git-example`.
-After entering the name, click "Create repository", and the next step is to link the new remote to the local repository.
-The required steps are detailed on screen when you do this at GitHub, but here we would execute the following command in the command line:
+Although we are creating a new repository here, the goal is to create an online "remote" repository for the running example within this tutorial.
+Because of this, do not check the "Initialize this repository with a README" (you already created one in the local repository), do not add a .gitignore or a license.
+For the current example, after clicking "New repository", we simply enter a name, which can be anything but for consistency we call the GitHub repository `git-example`---the same name as our local project folder.
+After entering the name, click "Create repository".
+The next step is to link the new remote to the local repository, and the required steps are detailed on screen when you do this at GitHub:
 
@@
 
-Where `<username>` and `<reponame>` should be replaced with the user's GitHub username, and the GitHub repository name.
+Where `<username>` and `<reponame>` are the user's GitHub username, and the GitHub repository name.
+The correct address is visible in the GitHub page that opens after creating the repository.
 Once the connection is set up, we can **push** local changes to the GitHub remote:
@@
 
-The `-u origin master` are only required for the first push, as they set up the connection. For pushing any changes collowing this, simply type `git push` after adding and committing locally.
-
-The team has now created the remote Central repository, and other users can start contributing to it.
+The `-u origin master` are only required for the first push, as they set up the connection.
+Running this command will send your local repository to the GitHub repository.
+For pushing changes collowing this initial push, simply type `git push` after adding and committing locally.
+You have now created the remote Central repository, and other users can start contributing to it.
 
@@
 Once another team member has set up Git on their own computer, and signed up for GitHub, they need to **clone** the remote GitHub repository onto their local computer.
-[TODO actually check that these steps are correct with a 'fresh' computer.]
-
-To clone a repository, the new user must first navigate to an appropriate location on the computer where they would like to create the repository on their computer, for example their `username/Documents` folder.
-Once the appropriate location is found, cloning will create a new folder for the repo inside this folder.
+To clone a repository, the new user must first navigate to an appropriate location on the computer where they would like to save the project on their computer, for example their `User/Documents` folder.
+Once the appropriate location is found, cloning will create a new folder for the repository inside this folder.
 
@@
 
-These new contributors can then work on their local copies as detailed in earlier parts of this tutorial; making changes, adding, and committing.
+To find out the correct link to enter to `git clone`, you can navigate your web browser to the repository's GitHub address (e.g.
+<https://github.com/mvuorre/reproguide-curate> for this tutorial's repository) and click the big green "Clone or download" button; the complete address is in the text box.
+
+New contributors can then work on their local copies as detailed in earlier parts of this tutorial; making changes, adding, committing, and pushing.
 After committing their changes, they can update the status of the Central Git repo by pushing their changes to it:
@@
 Just as you must manually push your own local changes to the remote repository, you must also obtain others' changes by **pulling** them from the central repo.
-Before starting to work on your proposed change, pull the remote changes with [TODO see if it needs to be rebased]:
+Pulling is indicated as the first step in the collaborative workflow in Figure \@ref(fig:git-diagram-collaborate), because it is important that you start working on the most up to date version of the project (e.g.
+you don't want to reinvent the wheel or make unnecessary conflicting changes).
+Before starting to work on your proposed changes, pull the remote changes with:
 
@@
 
-The way in which users, and their local repositories, interact with the central repository by pushing and pulling is the cornerstone of collaboration on GitHub, and thoughtful use of these commands allows for complex workflows without any important code (data, ideas in manuscript, analysis code) being ever overwritten.
-However, there is no automatic way for a computer to tell what changes to prioritize: If two users have worked on the same code, and they both attempt to push their changes, the user to do this later will have to resolve the **conflict**.
+The way in which users and their local repositories interact with the central repository by pushing and pulling is the cornerstone of collaboration on GitHub, and thoughtful use of these commands allows for complex workflows without any important code (data, ideas in manuscript, analysis code) ever being overwritten.
+However, there is no automatic way for a computer to tell what changes to prioritize: If two or more users have worked on the same code and then attempt to push their changes, it is possible that they have made changes that conflict with each other.
+If this happens, and it will, there is no need to worry; you simply need to know how to resolve the conflict.
 
@@
 
-If user A has pushed changes to the central repository while user B was working on the same code, user B's push will result in a merge conflict.
-That is just a natural consequence of two individuals working simultaneously on the same idea.
-User B simply needs to first merge the changes in the remote repository to her own code by pulling from the remote.
-Once the conflict has been manually resolved (and user B's code updated with the pull), user B can push changes from her local repository to the central repository.
+If user A has pushed changes to the central repository while user B was working on the same code, user B's later push can result in conflict.
+That is just a natural consequence of two individuals working simultaneously on the same idea, and then writing different code in the same location in the file.
+When this happens, user B needs to first pull the most up to date changes from the remote repository to her local code, and then inspect the code for problems (which will be apparent, Git will insert pointers to where the conflict is occurring.) Once user B's local code is updated with the pull, and the conflict manually resolved, user B can push changes from her local repository to the central repository.
+
 How these potential conflicts appear depends on how users collaborate with one another, and a detailed explanation of all potential scenarios is outside the scope of this tutorial.
@@
 The GitHub customer service is also very responsive to users' help requests.
+Most importantly, even in the event of merge conflicts, all committed changes are saved in Git's history and can be retrieved.
 
@@
 
-By default, all GitHub repositories are public, meaning that anyone with an internet connection can use their web browser to inspect the contents of your repository, or even clone it to their computer.
-This may sound intimidating to researchers who are used to working in a more private context, and clearly necessitates planning and thought with respect to issues such as data privacy and sharing sensitive materials.
+By default, all GitHub repositories are public: Anyone with an internet connection can use their web browser to inspect the contents of your repository, or even clone it to their computer.
+This may sound unfamiliar to researchers used to working more privately, and clearly necessitates planning and thought with respect to issues such as data privacy and sharing sensitive materials.
 However, for many projects---including the writing of this tutorial---we see very few downsides to working "in the open".
 
-There are two alternatives to working in a public GitHub repository: One, which we won't cover here, is to not use GitHub but instead place the central repository on the research team's shared server.
-The second option is to change the repository's settings on GitHub (this can be done when the repository is first created or afterwards) to private.
+There are two alternatives to working in a public GitHub repository: One, which we won't cover here, is to not use GitHub but instead place the central repository on the research team's shared but private server.
+The second option is to make the repository private on GitHub (this can be done when the repository is first created or afterwards by clicking Settings on the repository's website).
 Private repositories, and their contents, are only accessible to invited team members, and are therefore ideal for small teams who would like to work without revealing their master plans to the public just yet.
+For example, you might initially choose to work in a private repository, and only make it public once you feel the material is mature enough for public consumption.
 
-To make a GitHub repository private, navigate to the repository's website with a web browser (e.g.
-<https://github.com/<username>/<reponame>>), and click "Settings", then "Make this repository private".
-Once one user has set the central GitHub repository to as private, anyone wishing to clone, push, pull, or view the repo must provide their GitHub username and password.
+To make a GitHub repository private, navigate to the repository's website with a web browser, and click "Settings", then "Make this repository private".
+Once one user has set the central GitHub repository to private mode, anyone wishing to clone, push, pull, or view the repo must provide their GitHub username and password.
 Only if they match an invited team members username and password can the user access the repository.
@@
 [^tier-link]: <http://www.projecttier.org/tier-protocol/specifications/>
+[^git-branches]: A particularly powerful approach for trying new features is **branching**: The project can be duplicated to a new branch and modified, then merged back to the main branch after work on the new feature is complete---or the new branch can be discarded if the work ended unsatisfactory.
+Branches are outside the scope of this tutorial, for more information see the Git website (<https://git-scm.com/book/en/v2/Git-Branching-Branches-in-a-Nutshell>).
 

### 2017-06-16 manuscript/manuscript.Rmd
Edit  ## R Studio

@@
 
-Finally, we have above detailed how to use Git from the computer's command line, but not all users are comfortable---although there is no reason not to be!---with this mode of interacting with their computers.
-We chose to introduce Git from the command line because eventually users will need to use it for more complicated operations (and it is required for setting the user's information); however, there are various graphical user interfaces (GUIs) available for Git, that allow interacting with Git by pointing and clicking buttons with the mouse.
+Above, we have detailed how to use Git from the computer's command line, but not all users are comfortable---although there is no reason not to be!---with this mode of interacting with their computers.
+We chose to introduce Git from the command line because eventually users might want to use it for more complicated operations (and it is required for setting the user's information).
+However, there are various graphical user interfaces (GUIs) available for Git which allow point-and-click interacting with Git.
 
-In this section, we introduce how to use the popular R Studio Integrated Development Environment (IDE) for the R programming language for interacting with Git.
+## R Studio
+
+In this section, we introduce how to use the popular R Studio IDE [@rstudio_team_rstudio:_2016] for the R programming language [@r_core_team_r:_2017] for interacting with Git.
 An IDE is an interface that bundles together many necessary features of software development---some users may be familiar with R Studio for conducting statistical analyses.
@@
 
-For this example, we will begin a completely new repository using only R Studio, and assume that readers are familiar with the concepts from the above tutorial.
-
-## R Studio
-
-[TODO download and install R Studio, cite]
+For this example, we will begin a completely new project using only R Studio, and assume that readers are familiar with the concepts from the above tutorial.
+R Studio is a free and open source IDE, works on Windows, Mac, and Linux operating systems, and can be downloaded from the project's website at <https://www.rstudio.com/products/rstudio/download/>.
+Once downloaded and installed, it can be used for accessing the R programming language, and for many operations involved in curating research assets, including a GUI for Git.
 
@@
 
-[TODO put some screenshots in here]
+```{r rstudio-project-1, fig.cap = "Creating a project in R Studio."}
+knitr::include_graphics("images/rstudio-project-1.png")
+```
 

### 2017-06-16 manuscript/manuscript.Rmd
Add R Studio screenshots.

@@
 To start a new project with R Studio, open the R Studio application, and click File -> New Project.
-This brings up a dialog asking whether to create a project in a new directory, existing directory, or checkout an existing project from a version control repository.
-Here, we create a new project in a new directory, and choose "New Project" in the following screen.
+This brings up a dialog (left panel in Figure \@ref(fig:rstudio-project-1)) asking whether to create a project in a new directory, existing directory, or checkout an existing project from a version control repository.
+Here, we create a new project in a new directory, and choose "New Project" in the following screen (middle panel in Figure \@ref(fig:rstudio-project-1)).
 Then we'll give a name to the project's home folder and choose where to save it on the computer.
-Importantly, we'll also check the "Create a git repository" box, which will automatically set up a new repository for the project (provided you have set Git up as detailed above).
+Importantly, we'll also check the "Create a git repository" box (right panel in Figure \@ref(fig:rstudio-project-1)), which will automatically set up a new repository for the project (provided you have set Git up as detailed above).
 Clicking "Create project" creates the folder in the specified location, and two files inside the project's main folder.
 
-```{r rstudio-project-1, fig.cap = "Creating a project in R Studio."}
+```{r rstudio-project-1, fig.cap = "Creating a project in R Studio.", out.width="6in"}
 knitr::include_graphics("images/rstudio-project-1.png")

### 2017-06-16 manuscript/manuscript.Rmd
R Studio complete.

@@
 One of these files is `.gitignore` which we covered above.
-The other file is and .Rproj file, which indicates that the folder is the home folder for an R (Studio) project.
-Users don't interact with this file, but when opened, it is simply a plain text file containing the project's settings (these can be modified through Tools -> Project Options in R Studio).
+The other file is an `.Rproj` file, which indicates that the folder is the home folder for an R (Studio) project.
+Users don't interact with this file directly, but if opened, it is a plain text file containing the project's settings (these can be modified through Tools -> Project Options in R Studio).
 
@@
 
-To mark this milestone of creating a project, let's commit these changes to Git.
-The workflow is exactly the same as when using Git from the terminal, but we now have some visually appealing helpers.
+To mark this milestone of creating a project, commit all the changes so far (new files, new project) to Git.
+The workflow is exactly the same as when using Git from the terminal, but there are now some visually appealing helpers.
 To begin committing these changes, click on "Commit" in the Git tab.
-This brings up another window where we can select files to add into the staging area (the `.gitignore` file in Figure \@ref(fig:rstudio-git-2) is staged by checking the box).
-Once all the desired files (here we chose both) are staged, we write a short commit message as detailed above, and click the "Commit" button.
+This brings up another window where you can select files to add into the staging area (the files in Figure \@ref(fig:rstudio-git-2) are staged by checking the boxes on the names' left).
+Once all the desired files (here we chose both) are staged, you can write a short commit message in the appropriate text box, and click the "Commit" button.
+The text highlighted in green in the bottom part of Figure \@ref(fig:rstudio-git-2) indicates the lines of text that were changed in the selected file (`.gitignore`).
+These lines are all green because the file is new to Git and therefore each change is an addition (the files are unimportant R project files which users can safely ignore.)
 
@@
 When creating the R Project, there was an option for creating the project from an existing Git repository.
-Because we didn't do this, we must now manually instruct Git that this repo should have a remote in GitHub.
-This is done exactly as above, and we must use command line functions.
-After you've created a new repository on GitHub, navigate to the project's folder on your computer with the command line navigator of your choice, and execute the following commands:
+Because we didn't do this, we must now manually instruct Git that this project should have a remote GitHub repository.
+This is done exactly as above, using the command line functions.
+After you've created a new repository on GitHub, navigate to the project's folder with the command line shell (or you can click on "More" in R Studio's Git tab, and then "Shell", which opens a new shell window in the correct location), and execute the following commands (but replace <username> and <reponame> with the correct names):
 

### 2017-06-16 manuscript/manuscript.Rmd
Added discussion section

@@
 
-After executing these commands, you can use all Git and GitHub features from R Studio without having to use the command line.
-
-### Further examples
-
-- Born open data [@rouder_what_2015]
+After executing these commands, you can use all Git and GitHub features from R Studio.
+This is especially useful because the R Studio IDE offers a complete environment for project management, data analysis, and manuscript preparation with the R Markdown and knitr R packages [@allaire_rmarkdown:_2016; @xie_knitr:_2016].
+Psychologists will be especially interested in the papaja package for creating APA formatted manuscripts (@aust_papaja:_2016; the source code of this manuscript, which was prepared with the papaja package, can be found at <https://github.com/mvuorre/reproguide-curate>).
 
@@
 
-## Overview of Git Commands
+Table \@ref(tab:git-commands) gives the main Git commands and their purported use in roughly the order in which they would be used in a typical workflow.
+Help on how to use each command is available by appending `--help` to each command (e.g.
+`git commit --help`).
+When printed in the command line, some help pages run for several pages, press the spacebar to move to the next page, or `q` to quit looking at the help page.
+In Box 2 we give links to online materials that we have found particularly helpful in learning Git and teaching it to others.
 
@@
 \end{tcolorbox}
+
+# Discussion
+
+We have presented an introductory tutorial on using the Git version Control System for curating and collaborating on research assets in behavioral sciences.
+Although this tutorial includes enough material to get started and manage most operations, Git (and GitHub) is a vast ecosystem with endless opportunities.
+For example, the concept of "Born open data", where research data is automatically posted online upon collection, is made very easy with the Git + GitHub workflow [@rouder_what_2015].
+Another important feature of this workflow, which we did not cover, relates to disseminating scientific knowledge.
+The Git + GitHub workflow, especially when used from R Studio, makes it easy for researchers to create accessible documents in .pdf, website, blog post and other formats for facilitating communicating their research.
+
+Some limitations of using Git (and GitHub) are that it requires some initial learning up front, and human intervention at various steps when changes are committed, pushed, etc.
+Other common collaboration and "versioning" systems, such as Dropbox seem to manage these operations automatically and without requiring any time investment in learning the tools.
+Why, then, should researchers spend their valuable time in learning to use Git (or other VCSs) instead of alternatives (e.g.
+Dropbox).
+
+First, Dropbox doesn't save detailed information on who did what and when at each change point in a file's history: Although files' past versions are saved for 30 days (for free account users), it is impossible to tell what exactly changed between versions, and therefore finding the desired version of a file can be particularly difficult.
+With Git, you can easily `git log` back in time and see what happened at each change-point (especially if you took the time to write informative commit messages.)
+
+Second, we offer that the investment---which may itself be smaller than many might think---to learning Git can end up turning into saved time for more complex projects, because Git never loses track of a project's history, and allows for multiple collaborators to seamlessly work together on the same files.
+Collaborators don't need to give a file to one person to work on while all others wait for these changes.
+
+Third, the realities of scientific collaboration are too complex to easily narrow down to a model in which files are automatically updated across multiple computers, such as they are when using an "automatic" service like Dropbox.
+Sharing complex ideas and code across multiple collaborators will ultimately always require some form of human oversight and communication, and this insight has been recognized and implemented by software developers, and Git's tools for merging complex ideas from multiple sources.
+
+Finally, it is important to note that Version Control and backups of one's work are not the same thing.
+Although Git allows for keeping track of the history of one's project, it is not a physical or virtual backup of that work.
+Researchers should also maintain best practices in backing up their research assets in multiple physical locations, and possibly online as well.
+
+The challenges to reproducibility are many, and they have only recently received the targeted attention they deserve for the reliability of empirical sciences.
+Curating research assets and focusing on the scientific workflow is important for ensuring the continuity of one's work and improves efforts for a cumulative and reliable science.
+
 \newpage

### 2017-06-19 manuscript/manuscript.Rmd
Force knitting directory to project root; ignore error if file not found.

@@
 opts_chunk$set(
-    fig.align = "center"
+    fig.align = "center",
+    root.dir = "../"
 )
@@
 
-```{r git-commands, results = 'asis', warning = F}
+```{r git-commands, results = 'asis', warning = F, error=T}
 commands <- read_csv("data/git-commands.csv")

### 2017-06-19 manuscript/manuscript.Rmd
manuscript edits

@@
   - name          : "James P. Curley"
-    affiliation   : "1"
+    affiliation   : "1,2"
 affiliation:
@@
     institution   : "Department of Psychology, Columbia University, New York, USA"
+  - id            : "2"
+    institution   : "Department of Psychology, University of Texas at Austin, Texas, USA"
 
@@
 Reproducibility can be defined as follows: "A research project is computationally reproducible if a second investigator (including you in the future) can recreate the final reported results of the project, including key quantitative findings, tables, and figures, given only a set of files and written instructions." [@kitzes_practice_2017].
-In the context of experimental Psychology, for example, a "project" could be an experiment or a set of experiments investigating a hypothesis, "reported results" could be figures or statistical analyses in a conference presentation or a manuscript.
+In the context of Experimental Psychology, for example, a "project" could be an experiment or a set of experiments investigating a hypothesis, "reported results" could be figures or statistical analyses in a conference presentation or a manuscript.
 
@@
 
-Further, although they are sometimes erroneously conlated [@open_science_collaboration_estimating_2015; @peng_reproducible_2011] reproducibility and replicability are independent of one another.
+Further, although they are sometimes erroneously conflated [@open_science_collaboration_estimating_2015; @peng_reproducible_2011] reproducibility and replicability are independent of one another.
 Although the distinction between the two may be somewhat blurred in purely computational areas (e.g.
@@
 This note is usually called a readme, and therefore our first contribution to this project will be a README file (this file is so important that it has become standard practice to write it in capital letters).
-The README file should be a plain text file (i.e.
+The README file should be a plain text `.txt` file (i.e.
 not created with Microsoft Word) that can be read with a simple text editor.
-You can create this file with any text editor (Macs have TextEdit app installed by default, which allows creating plain text files).
+You can create this file with any text editor (e.g.
+the TextEdit and Notepad apps which allow for creating plain text files are installed by default on Mac and Windows OS respectively).
 We created the file with some text in it, and it can now be seen in the `git-example` folder, either by using the system file viewer or `ls` in the command line.
@@
 First, one user creates this central repository, then other users **clone** a local copy of it (see Table \@ref(tab:git-commands)).
-Contributors, including the one who created the central repository, then work on their local projects as detailed above, and after committing **push** their changes to the central repository.
+Contributors, including the one who created the central repository, then work on their local projects as detailed above, and after committing, **push** their changes to the central repository.
 To get changes that other users have pushed to the central repository, contributors **pull** changes from it.
@@
 [^git-link-submodules]: For advanced users, Git *submodules* allow linking projects to each other, or organizing more complex projects into projects and their sub-projects (<https://git-scm.com/book/en/v2/Git-Tools-Submodules>).
-[^git-cheat-sheet]: It helps to have a Git command cheat sheet (<https://services.github.com/resources/cheatsheets/>) printed and taped on your wall, but it contains many more commands than are needed for the basic use of Git.
+[^git-cheat-sheet]: It helps to have a Git command cheat sheet (<https://services.github.com/resources/cheatsheets/>) printed and taped on your wall, but it contains many more commands than are needed for the basic use of Git in standard Psychology studies.
 [^tier-link]: <http://www.projecttier.org/tier-protocol/specifications/>

### 2017-06-19 manuscript/manuscript.Rmd
edit typos

@@
 
-Given this broad definition of reproducibility, it is easy to recognize that many projects fall between the extremes of completely replicable and not replicable at all.
+Given this broad definition of reproducibility, it is easy to recognize that many projects fall between the extremes of completely reproducible and not reproducible at all.
 A completely reproducible project would provide consumers the entire raw data set, and detailed, unambiguous instructions to analyze it (usually as computer code).

### 2017-06-19 manuscript/manuscript.Rmd
Everything is code; contrast VCS to "standard workflow"

@@
 Given this broad definition of reproducibility, it is easy to recognize that many projects fall between the extremes of completely reproducible and not reproducible at all.
-A completely reproducible project would provide consumers the entire raw data set, and detailed, unambiguous instructions to analyze it (usually as computer code).
+A completely reproducible project would provide consumers the entire raw data set with detailed and unambiguous instructions to analyze it (usually as computer code).
 This degree of reproducibility is very rare in Psychology [@wicherts_poor_2006; @vanpaemel_are_2015], although data sharing has become more common in recent years in journals that award authors who do so with special badges [@kidwell_badges_2016].
@@
 
-Version Control Systems are computer programs designed for tracking changes to computer code, and collaborating on the code with others.
-Although initially developed for writing code collaboratively[^vcs-collaborate], it is easy to recognize that VCS can also be adopted for scientific production and collaboration.
-For example, behavioral experiments are often created with a programming language, and can have multiple authors and versions.
+Version Control Systems (VCS) are computer programs designed for tracking changes to computer code, and collaborating on the code with others.
+VCS were initially developed by software engineers for writing code collaboratively[^vcs-collaborate], but are increasingly being adopted to enhance workflows outside computer science.
+To help understand why, it is helpful to think of "code" more broadly as any text written on a computer: Manuscripts, books, statistical analysis scripts, source code of computerized experiments, and even data files are "code" or have source code, which is just plain text written on a computer, and which can benefit from version control.
+Going forward, when we use the word "code" in this tutorial, we mean it in this broad sense (e.g.
+this manuscript's source "code" was written on a computer, and was version controlled.)
+
+For example, computerized behavioral experiments are written with a computer programming language such as MATLAB or Python.
+The experiment's source code (text written by human, interpreted by computer to e.g.
+display stimuli to participants) may have multiple authors and usually goes through multiple versions.
 Keeping track of the versions of the program and changes to the code, and allowing many authors to contribute to it (without breaking the experimental program) are problems that VCSs were specifically designed for.
-VCS is also easily adopted to other aspects of the research cycle, such as curating data across computers and laboratories, or writing manuscripts with multiple authors, versions, and sources of results.
+It might be less obvious that writing a manuscript is quite similar, as far as the computer is concerned: Multiple authors write multiple versions of a text document, and sometimes previous versions need to be inspected, and the text needs to be "merged" across the many authors.
+Even less obvious is that data itself is often plain text: In most computerized behavioral experiments, the output data are numbers and text written into a text file.
+These text files can then be version controlled (and researchers would usually not want to see that their raw data files have changed after they were created---VCS allows verifying that they haven't changed because their history is logged.)
 
@@
 
-Some popular Version Control Systems are SVN, Mercurial, and Git.
-In this tutorial, we focus on Git: Git is already incresing in popularity in the scientific community, and is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.
+To understand what VCS is, it is useful to contrast an example common case with VCS and what we believe is a fairly standard workflow.
+Consider collaborating with one person on a manuscript that reports results from a data set using some statistical model.
+In the standard workflow, one person might organize the raw data in one way to fit a statistical model, and then write a draft of the manuscript.
+This would create three files: `data.csv`[^csv], `analysis.R`, and `manuscript.doc`.
+If the coauthor then wished to explore another statistical model, which requires the data in a different format, and then edit the manuscript, she would create three more files: `data_new.csv`[^csv], `analysis_new.R`, and `manuscript_new.doc`.
+This cycle would then repeat as many times as required, each time creating more files, making it more and more difficult for the authors to exactly remember which data was paired with which analysis, and which analyses (and data format) was reported in which version of the paper.
+
+Using VCS the workflow is greatly simplified: There are only three files (data, analysis, and manuscript), but they are all under version control, which keeps track of changes to them.
+Because changes are monitored, creating new files for every new idea or edit is unnecessary, thereby possibly reducing room for error in remembering which data file was linked with which analysis, and which manuscript version had the correct numerical results, and so forth.
+Further, as we explain below, VCS allows many coauthors to work on the files simultaneously, eschewing the need for emailing a version to a coauthor, then waiting for them to make changes before continuing with your work.
+Box 1 has further practical examples of where VCS can be used to improve the scientific workflow.
 
@@
 ## The Git Version Control System
+
+Version control software has a long history in software engineering, and there are many VCS programs.
+Some popular ones are Apache Subversion[^subversion], Mercurial[^mercurial], and Git.
+In this tutorial, we focus on Git: Git is already incresing in popularity in the scientific community, and is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.
 
@@
 
+[^csv]: `.csv` stands for comma separated values, and is a plain text file commonly used to store two-dimensional data.
 [^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
@@
 Branches are outside the scope of this tutorial, for more information see the Git website (<https://git-scm.com/book/en/v2/Git-Branching-Branches-in-a-Nutshell>).
+[^subversion]: <https://subversion.apache.org/>
+[^mercurial]: <https://www.mercurial-scm.org/>
 

### 2017-06-19 manuscript/manuscript.Rmd
Clarify install instructions
Add command line interface intro incl. navigating folders

@@
 For Mac users, the easiest way to install or update Git to the latest version is to download the installer from <http://git-scm.com/download/mac>, and install it like any other program.
-Windows users can download the Git software installer from <http://git-scm.com/download/win>.
+Similarly, Windows users can download the Git software installer from <http://git-scm.com/download/win> and install it like any other application.
 Linux (and other) users can download Git and find further install instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>).
@@
 
-Once you have installed Git, you can be use it from the command line or through a GUI--there are various GUIs for using Git, and below we will explain how to use the R Studio Git GUI.
-First, it is important to learn a few basic Git commands from the OSs command line.
+Once you have installed Git, you can use it from the computer's command line or through a GUI---there are various GUIs for using Git, and below we will explain how to use the R Studio Git GUI.
+First, it is important to learn a few basic Git commands from the OSs command line, because it is simultaneously the most basic and most flexible interface to Git's functionality.
+Understanding the basic Git commands as entered through the computer's command line facilitates understanding its more advanced uses, and is highly recommended, hence we begin using Git from the command line.
+
+### The command line
+
 The command line is a text-based interface for interacting with your computer, and its functionality greatly extends that of the standard way of interacting with the computer by clicking and pointing with a mouse.
+Many advanced techniques require using your computer through a command line, and it is very helpful in e.g.
+scripting and scheduling tasks on your computer.
+Here, we introduce the command line in just enough detail so that you can navigate folders on the computer, and set up Git's basic configuration (identify Git's user).
 
 To access the command line interface, you need to use a command line "shell" application.
-Mac users can open the built-in app *Terminal*, and Windows users can download and use one of many alternatives, such as Microsoft PowerShell (<https://msdn.microsoft.com/powershell>).
-After opening the shell app, you can type in commands to access the interface.
-For example typing `ls` (hit Return to execute) lists the files in the current folder.
-Some users may find using a text-based command line interface unfamiliar, but to get started with Git, there are only two required configuration commands which you need to run once, and the basic functionality requires using only four commands[^cli-links].
+Mac users can open the built-in app *Terminal*, and Windows users can use Git Bash, which is provided with the Git for Windows program.
+After opening the shell app, you can type in commands and execute them by pressing Return (Mac) or Enter (Windows).
+
+For example typing and executing `ls` in the command line lists the files in the current folder.
+First, you will want to know how to navigate the folders on your computer (a task that is typically done by clicking folders in Finder (Mac) or File Explorer (Windows)).
+Typically, when you open up your command line shell, you begin in the user's home directory.
+Depending on your operating system, the home directory is often represented with a `~` on the left side of the cursor.
+To ask for the current working directory, you can use the function `pwd`.
+To move up in the directory hierarchy (into the folder that contains the current directory), you can use `cd ..` (note the space).
+To move into a folder that is inside the current working directory, you can use `cd folder` where `folder` is the name of the desired folder.
+These command line functions are illustrated in Figure \@ref(fig:cli-intro).
+
+```{r cli-intro, fig.cap="The Terminal command line shell.
+The functions are explained in more detail in the text.
+You command line interface might look slightly different because of different user names, operating systems, and command line shell applications."}
+knitr::include_graphics("images/terminal-commands.png")
+```
+
+Some users may find using a text-based command line interface unfamiliar, but to get started with Git, there are two required configuration commands which you need to run once, and the basic functionality requires using only four commands[^cli-links].
 However, users can also read through the first part of the tutorial without executing the commands and simply focus on the workflow; we show how to use Git with a GUI in the second part of the tutorial.
+
+### Setting Git user info
 

### 2017-06-19 manuscript/manuscript.Rmd
Incorporating comments & edits

@@
 Each Git command starts with the word `git`, then a command (such as `config`), and then arguments to the command, such as `--global` (for global configuration) followed by the `user.name` variable name to show the global user name.
-For detailed instructions on how to use Git commands, you can type `git --help`.
-For help on how to use the `git config` command, type `git config --help`.
+
+```{r git-setup, fig.cap="Setting up Git's configuration using the command line.
+Notice that for these configuration commands, the current working directory does not matter (we did them in the user's home directory, indicated by \~)."}
+knitr::include_graphics("images/git-setup.png")
+```
 
@@
 You can verify that the correct email address was saved by typing `git config --global user.email` in the Terminal, and hit return.
+These commands are illustrated in Figure \@ref(fig:cli-intro).
 
 Once this information is entered, Git will know who you are, and is able to track who is doing what and when within a project, which is especially helpful when you are collaborating with other people, or when you are working on multiple computers.
+For detailed instructions on how to use Git commands, you can type `git --help`.
+For help on how to use the `git config` command, type `git config --help`.
 
@@
 The first operating principle of Git is that your work is organized into independent projects, which Git calls *repositories*[^git-link-submodules].
-A repository is a folder on your computer which is version controlled by Git (you can tell if a folder is "monitored" by git by checking if the folder, or any of its parent folders, contains a hidden `.git` directory).
+A repository is a folder on your computer which is version controlled by Git[^git-folder].
 Everything that happens inside a repository is tracked by Git, but you have full control of what is committed to Git's history and when.
@@
 
-```{r git-diagram, fig.cap = "A diagram illustrating the typical Git workflow. Verbs indicated in bold text are Git operations and explained in detail in the main text."}
+```{r git-diagram, fig.cap = "A diagram illustrating the typical Git workflow.
+Verbs indicated in bold text are Git operations and explained in detail in the main text.
+In brief, the typical Git workflow consists of making changes to files in the project (such as editing a manuscript.) Once the revised version of the file is saved, the user adds the changed file to the staging area.
+Many files can be added to the staging area, if desired.
+Once the changes in the staging area reflect a conceptual entity, such as edits to an image file and its accompanying caption in the manuscript, the user then commits the changes to Git's history.
+These commits are accompanied with short commit messages that describe the changes made in that commit.
+Finally, when needed, changes to files can be discarded by checking out an earlier version of the file from Git's history."}
 knitr::include_graphics("../data/git-diagram.png")
@@
 It is therefore important to organize the project keeping a few key goals in mind (here, we follow guidelines such as the Project TIER recommendations[^tier-link]): The files and folders should have easy to understand names (avoid idiosynchratic naming schemes), and the names should indicate the purposes of the files and folders.
+We illustrate some good and not-so-good practices in Figure \@ref(fig:organizing-files).
+
+```{r organizing-files, fig.cap = "Organizing one's project helps potential collaborators (including oneself in the future) to quickly and reliably find the correct files.
+Some bad practices (left panel) include having multiple versions of each file, increasing the potential for e.g.
+accidentally using the wrong data for the analysis, or forgetting which version was used in the analysis.
+Better practices (right panel) organize the data to subfolders with meaningful names, and have one file per purpose.
+These files are then versioned using Git, eschewing the need for multiple files, thereby reducing potential for mistakes."}
+knitr::include_graphics("images/git-diagram.png")
+```
 
@@
 All the materials related to this project will be placed in subfolders of the home directory, but for now `git-example` is just an empty folder waiting to be filled with data, code, and documents.
-To make the example concrete, you can follow the tutorial by typing out the following commands on your own computer.
+To make the example concrete, we recommend that you follow the tutorial along on your own computer.
 
@@
 First, choose an appropriate folder on your computer (such as `User/Documents/`) where you'd like to create the project.
-You can either use the system's file navigator (Finder / File Explorer) to create this folder, or the following commands in the command line: Navigate to the desired folder by using `cd Documents` to move into the `Documents` folder (assuming it exists in the folder where you currently are), and `cd ..` to move out of the folder (to its containing folder.) Once you are in the folder where you want to create the project, type `mkdir git-example` to make the `git-example` directory, then `cd git-example` to move into it.
-Once you are in the project's home folder (you can verify where you are by typing `pwd`), you want to turn the folder into a project by initializing Git:
+You can either use the system's file navigator (Finder / File Explorer) to create this folder, or use the command line: Navigate to the desired folder by using `cd Documents` to move into the `Documents` folder (assuming it exists in the folder where you currently are).
+Use `cd ..` to move out of a folder (to its containing folder), if needed.
+Once you are in the folder where you want to create the project, type `mkdir git-example` to make the `git-example` directory, then `cd git-example` to move into it.
+Once you are in the project's home folder (you can verify where you are by typing `pwd`), you want to turn the folder into a project by initializing Git with the `git init` command.
+These commands are shown in \@ref(fig:initialize-git)
+
+```{r git-initialize, fig.cap = "Creating and navigating to a folder, and initializing it as a Git repository."}
+knitr::include_graphics("images/git-diagram.png")
+```
+
+Instead of screenshots, for the rest of the tutorial we present the commands as follows:
 
 ```bash
-> git init
+$ git init
 ```
 
-This command initializes the folder as a Git repository, and the only change so far has been the addition of a hidden `.git` folder inside `git-example` (and a `.gitattributes` file.
-Users can ignore these hidden files and folders, however they can be shown by typing `ls -la`.) Now that the folder is initialized as a Git repository, Git monitors any changes within it, and allows you to add and commit these changes (Figure \@ref(fig:git-diagram).)
+Where each command is preceded by a `$` symbol.
+To help remember the commands, we recommend simply typing them out, instead of copy-pasting.
+The `git init` command initializes the folder as a Git repository, and the only change so far has been the addition of a hidden `.git` folder inside `git-example` (and possibly a `.gitattributes` file.
+Users can ignore these hidden files and folders, however they can be shown by the `ls -la` command.) Now that the folder is initialized as a Git repository, Git monitors any changes within it, and allows you to add and commit these changes (Figure \@ref(fig:git-diagram).)
 
@@
 The README file should be a plain text `.txt` file (i.e.
-not created with Microsoft Word) that can be read with a simple text editor.
+not created with Microsoft Word) that can be read with a simple text editor[^plain-text].
 You can create this file with any text editor (e.g.
@@
 Because we added this file to a Git repository, Git is also aware of it.
-To see what files have changed since the last status change in the repository (there clearly has been only this one), you can ask for Git's **status**:
+To see what files have changed since the last status change in the repository (there clearly has been only this one), you can ask for Git's **status**.
+We present commands prepended with `$` symbols, and output printed without preceding characters:
 
 ```bash
-> git status
+$ git status
 
@@
 ```bash
-> git add README
+$ git add README
 ```
@@
 ```bash
-> git commit -m "Add README file."
+$ git commit -m "Add README file."
 ```
@@
 ```bash
-> git status
+$ git status
 
@@
 It shows changes within `<file>`, line by line, highlighting removals with red and added lines with green.
-Once you are happy with the changes, you can repeat the add and commit steps from above to permanently record the current state of the project to Git's log (below we use the `.` shortcut for adding all changes):
+Once you are happy with the changes, you can repeat the add and commit steps from above to permanently record the current state of the project to Git's log (below we use the `.` shortcut for adding all changes[^dot-shortcut]):
 
 ```bash
-> git add .
-> git commit -m "Populate README with project description."
+$ git add .
+$ git commit -m "Populate README with project description."
 ```
@@
 ```bash
-> git log
+$ git log
 ```
@@
 ```bash
-> git show 60cbe5c9b4a78e500314f791080381030577a035
+$ git show 60cbe5c9b4a78e500314f791080381030577a035
 commit 60cbe5c9b4a78e500314f791080381030577a035
@@
 ```bash
-git add .gitignore
-git commit -m "Added .gitignore file"
+$ git add .gitignore
+$ git commit -m "Added .gitignore file"
 
-git add .
-git commit -m "Completed power analysis"
+$ git add .
+$ git commit -m "Completed power analysis"
 ```
@@
 ```bash
-git checkout experiment-1/analysis/plan_n.R
+$ git checkout experiment-1/analysis/plan_n.R
 ```
@@
 ```bash
-git remote add origin https://github.com/<username>/<reponame>.git
+$ git remote add origin https://github.com/<username>/<reponame>.git
 ```
@@
 ```bash
-git push -u origin master
+$ git push -u origin master
 ```
@@
 ```bash
-git clone https://github.com/<username>/<reponame>.git
+$ git clone https://github.com/<username>/<reponame>.git
 ```
@@
 ```bash
-git push
+$ git push
 ```
@@
 ```bash
-git pull
+$ git pull
 ```
@@
 ```bash
-git remote add origin https://github.com/<username>/<reponame>.git
-git push -u origin master
+$ git remote add origin https://github.com/<username>/<reponame>.git
+$ git push -u origin master
 ```
@@
 [^mercurial]: <https://www.mercurial-scm.org/>
+[^git-folder]: There are no visible changes to a folder once it is tracked by Git.
+Once Git is initialized in a folder, the only change is that a hidden folder, called `.git` is added, but users do not need to interact with it directly.
+[^plain-text]: Plain text has many advantages over proprietary file formats such as Microsoft Word's .docx files.
+Briefly, plain text is both human and computer readable (.docx are unreadable without a copy of Microsoft Word), is both forward and backward compatible (there will always, and has always been, software capable of reading it), and plain text takes very little space.
+The file extension of plain text doesn't matter much, but we recommend using either `.txt` because it is widely recognized, `.md` for markdown syntax, or in the case of a readme file, simply not using any file extension.
+[^dot-shortcut]: For the `.` shortcut to work, you must currently be in the project's root directory.
+If you are not in the root directory, you can use `-A` shortcut instead; however, we recommend that you always ensure that you are in the project's root directory when running any Git commands.
 

### 2017-06-19 manuscript/manuscript.Rmd
Replace code folder tree with screenshot

@@
 ```{r git-setup, fig.cap="Setting up Git's configuration using the command line.
-Notice that for these configuration commands, the current working directory does not matter (we did them in the user's home directory, indicated by \~)."}
+Notice that for these configuration commands, the current working directory does not matter (we did them in the user's home directory, indicated by \\~)."}
 knitr::include_graphics("images/git-setup.png")
@@
 These files are then versioned using Git, eschewing the need for multiple files, thereby reducing potential for mistakes."}
-knitr::include_graphics("images/git-diagram.png")
+knitr::include_graphics("images/project-organization.png")
 ```
@@
 ```{r git-initialize, fig.cap = "Creating and navigating to a folder, and initializing it as a Git repository."}
-knitr::include_graphics("images/git-diagram.png")
+knitr::include_graphics("images/git-initialize.png")
 ```
@@
 
-For example, after working for a while on the project, you could have added two files inside well-organized subfolders of the project. The project would like this when viewed with a file explorer:
+For example, after working for a while on the project, you could have added two files inside well-organized subfolders of the project.
+The project would like Figure \@ref(fig:advanced-project) when viewed with a file explorer:
 
-```bash
-git-example/
-|-- admin/
-    |-- ethics-info.pdf
-|-- experiment-1/
-    |-- analysis/
-        |-- plan_n.R
-|-- manuscript/
-|-- README
+```{r advanced-project, fig.cap = "A project with more components."}
+knitr::include_graphics("images/advanced-project.png")
 ```

### 2017-06-20 manuscript/manuscript.Rmd
Improve writing. Add note on when to commit.

@@
 For example, after working for a while on the project, you could have added two files inside well-organized subfolders of the project.
-The project would like Figure \@ref(fig:advanced-project) when viewed with a file explorer:
+The project would like Figure \@ref(fig:advanced-project) when viewed with a file explorer.
 
@@
 You would probably like to track any changes to the power analysis script, but the ethics information file isn't really something that you need to keep track of---at least for this example.
-It would be laborious to constantly keep ignoring it when committing changes to Git (especially if we'd like to commit multiple files simultaneously with `git add .`) Git has an elegant solution to specifying which files to keep track of.
+It would be laborious to constantly keep ignoring it when committing changes to Git (especially if you'd like to add multiple files simultaneously with `git add .`) Git has an elegant solution to specifying which files to keep track of.
 Because by default all files are monitored, Git uses a special file for instructing which files are to be *ignored*.
@@
 In the current example you could make Git ignore the `admin/` folder entirely, and any file with the .pdf extension inside `manuscript/`.
-The example .gitignore file would look like this (lines beginning with `#` are comments):
+The example `.gitignore` file would look like this (lines beginning with `#` are comments, and only read by humans):
 
@@
 
-Re-running `git status` now only shows the `plan_n.R` file (and the newly created .gitignore file, which is also under version control, naturally.)
+Re-running `git status` now only shows the `plan_n.R` file (and the newly created `.gitignore` file, which is also under version control, naturally.)
 
-Because there are now two untracked files that are not in .gitignore (`.gitignore` and `plan_n.R`), and you usually should aim to maintain a clean history for the project, you can create two separate commits: One for the Git ignore file, and one for the power analysis file.
+Because there are now two untracked files that are not specified to be ignored in `.gitignore` (`.gitignore` and `plan_n.R`), and you usually should aim to maintain a clean commit history for the project, you can create two separate commits: One for the `.gitignore` file, and one for the power analysis file.[^when-to-commit]
 
@@
 
-After this last commit, you can at any time come back to this commit with `git log` or `git show` and see what was inside the newly created power analysis file when it was first created.
+After this last commit you can, at any time in the future, come back to this commit with `git log` or `git show` and see what was inside the newly created power analysis file when it was first created.
 For instance, if new information suggests that you should change the assumed effect size in the power analysis, you can simply edit and save the file, then add and commit the changes to Git with a helpful message that logs this important event in Git's history.
@@
 If you are not in the root directory, you can use `-A` shortcut instead; however, we recommend that you always ensure that you are in the project's root directory when running any Git commands.
+[^when-to-commit]: It is entirely up to the user to decide what to commit and when.
+However, it is best practice to commit often while making incremental changes.
+Each commit should aim to solve one problem, introduce one new idea, or---more generally---do one thing.
+This way, when the commit history is reviewed later, it is easy to find and come back to a specific change.
 

### 2017-06-20 manuscript/manuscript.Rmd
Remove unnecessary whitespace in code examples.

--- 2017-06-19 manuscript/manuscript.Rmd
@@
 admin/
-
 # Ignore all .pdf files in the 'manuscript' folder
 manuscript/*.pdf
-
 # Ignore a specific file in a specific folder
@@
 $ git commit -m "Added .gitignore file"
-
 $ git add .

### 2017-06-20 manuscript/manuscript.Rmd
Add sections on how to revert to earlier file or earlier version of project.

@@
 All the materials related to this project will be placed in subfolders of the home directory, but for now `git-example` is just an empty folder waiting to be filled with data, code, and documents.
+Because the folder structure on a computer is easy to think of as a tree, a project's home folder---or any folder that has subfolders---is also known as the *root* directory.
+In what follows we use the terms home directory and root (directory) interchangeably.
 To make the example concrete, we recommend that you follow the tutorial along on your own computer.
@@
 With these example operations, we have discussed the main Git operations as outlined in Figure \@ref(fig:git-diagram): Make changes to files, **add** to staging area, **commit** to history; **checkout** from history to undo changes.
+
+### Undoing committed changes
+
+Another common scenario is that a user makes changes to a file, adds the changes to the staging area, and commits them to Git's history, and only then realizes that the changes weren't good.
+If you have committed changes to a file, and would like to revert back to an older version of the file, you could **checkout** the file's earlier version, and then commit the older version.
+[^undoing-more-info] For example, suppose you have made bad changes to a file called `file.txt`, and committed the changes to history, and would then like to undo the bad changes by reverting to an older version of the file.
+First, view the history with `git log` (here we use the `--oneline` argument to make the output more concise):
+
+```bash
+$ git log --oneline
+4c64f11 Bad changes to file.txt
+039d6ff Good changes to file.txt
+a73f2ec Add file.txt
+```
+
+Recall in `git log` returns most recent changes at the top, and notice that the `--oneline` argument has also made the commit hash codes shorter and thus easier to read and write.
+Here we can see that commit `039d6ff` has a good version of the file, and subsequent changes in commit `4c64f11` were bad (you would of course not commit "bad changes", but here the message is informative for clarity).
+To revert `file.txt` to it's good state in commit `039d6ff`, you can use use `git checkout <hash> <filename>`, which here would be:
+
+```bash
+$ git checkout 039d6ff file.txt
+```
+
+Now asking for `git status` reveals that `file.txt` has been modified in the working directory (its current state is as it was in the good commit `039d6ff`).
+You can now add and commit these changes with `git add file.txt`, then `git commit -m "Undid bad changes to file.txt"` (type a commit message suitable for your situation).[^undo-the-revert] `git log --oneline` would then show:
+
+```bash
+$ git log --oneline
+bcbb123 Undid bad changes to file.txt
+4c64f11 Bad changes to file.txt
+039d6ff Good changes to file.txt
+a73f2ec Add file.txt
+```
+
+This operation of checking out earlier versions is very useful not only for undoing changes, but for viewing older versions of files as well.
+However, if you would only like to view past states of the project, instead of reverting / undoing to an earlier state, you can checkout an earlier version of the repository, as explained below.
+
+### Going to an earlier version of the project
+
+To return to an earlier state of the project, you can use the `git checkout` command.
+For example, the display above shows that in commit `a73f2ec`, you had added `file.txt`.
+If you would like to see the project at that commit, type `git checkout a73f2ec`.
+This command instantly checks out the file(s) at that point in history, and you can view them.
+This is very helpful if, for example, you would like to quickly run an earlier version of a statistical analysis.
+After you have viewed this old version of the project, you can return to the current version with `git checkout master`.
+
+Neither of these operations---checking out an earlier version of a file or of the entire project---affects the history of the project, and are therefore "safe", in the sense that your project's history won't be overwritten.
+However, checking out an earlier version of a specific file changes the current state of the project (the current version of the file is temporarily overwritten with the old version), so it is good practice to carefully keep track of the current version of your file before making further commits.
 
@@
 This way, when the commit history is reviewed later, it is easy to find and come back to a specific change.
+[^undoing-more-info]: For more information on undoing changes, see <https://www.atlassian.com/git/tutorials/undoing-changes>.
+[^undo-the-revert]: If, for some reason, you preferred the latest version after all, you can undo the revert process by `git checkout HEAD file.txt`, instead of adding and committing the older version.
+This function checks out the current state ("HEAD") which contained the bad changes.
 

### 2017-06-20 manuscript/manuscript.Rmd
Complex sentences are less fun that simple ones.

@@
 
-Neither of these operations---checking out an earlier version of a file or of the entire project---affects the history of the project, and are therefore "safe", in the sense that your project's history won't be overwritten.
+Both of these operations---checking out an earlier version of a file or of the entire project---are "safe" in the sense that your project's history won't be affected.
 However, checking out an earlier version of a specific file changes the current state of the project (the current version of the file is temporarily overwritten with the old version), so it is good practice to carefully keep track of the current version of your file before making further commits.
